---
title: "Climate Water Loss Experiment - Capture Hydration Analysis"
author: "Savannah Weaver"
date: "2021"
output: pdf_document
toc: TRUE
---


# Packages

```{r setup, include = TRUE, message = FALSE}
if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse") # workflow and plots
if (!require("zoo")) install.packages("zoo")
library("zoo") # interpolation using na.approx
if (!require("weathermetrics")) install.packages("weathermetrics")
library("weathermetrics") # F to C conversion
if (!require("lme4")) install.packages("lme4")
library("lme4") # for LMMs
if (!require("lmerTest")) install.packages("lmerTest")
library("lmerTest") # for p-values
if (!require("UsingR")) install.packages("UsingR")
library("UsingR") # simple.eda model assumption checker
if (!require("ggpubr")) install.packages("ggpubr")
library("ggpubr") # for multi-ggplot figs
if (!require("broom")) install.packages("broom")
library("broom") # lmer model export
if (!require("broom.mixed")) install.packages("broom.mixed")
library("broom.mixed") # lmer model export
if (!require("AICcmodavg")) install.packages("AICcmodavg")
library("AICcmodavg") # model selection
if (!require("car")) install.packages("car")
library("car") # VIFs
```


# Background and Goals

This data was collected June - August by Master's student Savannah Weaver, advisor Dr. Emily Taylor, and research assistants Tess McIntyre and Taylor Van Rossum. Adult male *Sceloporus occidentalis* were caught across the Cal Poly campus and in Poly Canyon. This R file analyzes the state and variation of osmotic balance and regulation at the time of capture. Please refer to **doi:** for the published scientific journal article and full details.


# Data

## Load

Read-in and attach all data. Details described later.

```{r load all lizard data}
        # mass and hematocrit data
full_dat <- read.csv("./data/mass_hct_notes.csv", # filename
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
        # format date to enable joining by date
  mutate(measurement_date = as.character(as.Date(measurement_date, 
                                          format = "%m/%d/%y"))
         ) %>%
        # join plasma osmolality data
  left_join(read.csv("./data/osml_means_clean.csv", # filename
                             na.strings=c("","NA") # fix empty cells
            ), by = c("individual_ID", 
                      "measurement_date" = "date_blood_drawn")
            ) %>%
        # join CEWL data
  left_join(read.csv("./data/CEWL_dat_all_clean.csv", # filename
                             na.strings=c("","NA") # fix empty cells
            ), by = c("individual_ID", 
                      "measurement_date" = "date")
            ) %>%
        # select variables of interest only
  dplyr::select(measurement_date,
                time_captured,
                time_processed,
                time_c_temp,
                type, day, 
                individual_ID,
                mass_g, 
                hemolyzed,
                hematocrit_percent, 
                osmolality_mmol_kg_mean,
                CEWL_g_m2h_mean = CEWL_g_m2h,
                msmt_temp_C,
                msmt_RH_percent,
                cloacal_temp_C
                ) %>%
          # format date-time-related variables
  mutate(measurement_date = as.Date(measurement_date, 
                                    format = "%Y-%m-%d"),
         # compute vapor pressure deficit
         msmt_temp_K = msmt_temp_C + 273.15,
         # find saturation level first
         e_s_kPa_m = 0.611*exp((2500000/461.5)*
                                  ((1/273)-(1/msmt_temp_K))),
         # actual vapor pressure
         e_a_kPa_m = e_s_kPa_m * (msmt_RH_percent/100),
         # VPD
         msmt_VPD_kPa = e_s_kPa_m - e_a_kPa_m
         ) %>%
  group_by(individual_ID) %>%
          # for each individual, extract capture date
  mutate(capture_date = min(measurement_date),
         day_n = as.numeric(measurement_date - capture_date))

summary(full_dat)
# check
unique(full_dat$capture_date)
```


## Export

Export full_dat to be used in 'experiment_analysis'.

```{r export exp data}
write.csv(full_dat, "./data/full_exp_data.csv")
```






## Format

Extract only the data from capture day (1 row of observations for each individual) and format the data classes properly for analysis.

```{r format cap data}
capture_dat <- full_dat %>%
          # select only data from capture days
  dplyr::filter(day_n == 0) %>%
  left_join(read.csv("./data/tmt_assignments.csv"),
            by = "individual_ID") %>%
          # put date and time together
  mutate(capture_date_time = (paste(capture_date, time_captured)),
         capture_date_time = as.POSIXct(capture_date_time, 
                                        format = "%Y-%m-%d %H:%M"),
          # correctly format time-only variables
         time_captured = as.POSIXct(time_captured, 
                                     format = "%H:%M"),
         time_processed = as.POSIXct(time_processed, 
                                      format = "%H:%M"),
         time_c_temp = as.POSIXct(substr(time_c_temp, 12, 16), 
                                  format = "%H:%M"),
          # set categorical variables as factors
         type = as.factor(type),
         day = as.factor(day),
         individual_ID = as.factor(individual_ID),
         hemolyzed = as.factor(hemolyzed),
          # set numeric measurements as numeric
         mass_g = as.numeric(mass_g),
         hematocrit_percent = as.numeric(hematocrit_percent),
         osmolality_mmol_kg_mean = as.numeric(osmolality_mmol_kg_mean),
         CEWL_g_m2h_mean = as.numeric(CEWL_g_m2h_mean),
         cloacal_temp_C = as.numeric(cloacal_temp_C)
                ) %>%
  # make sure only complete data included
  # this removes the data for individuals 304 (recapture) & 254 (escapee)
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean,
                               CEWL_g_m2h_mean, cloacal_temp_C)) %>%
  # remove experiment variables not relevant to capture analysis
  dplyr::select(-trial_number, -temp_tmt, -humidity_tmt, 
                -conclusion, -notes, 
                -shed, -tail_broken, -died)
summary(capture_dat)
```




Also do some summary stats for permit reporting:

```{r}
step1 <- full_dat %>%
  group_by(individual_ID) %>%
  summarize(capture_date = mean(capture_date))
  
permit_stats <- capture_dat %>%
  group_by(capture_date) %>%
  summarise(n = n()) %>%
  mutate(sex = "M")
permit_stats
sum(permit_stats$n)
write.csv(permit_stats, "./data/collection_summary.csv")
```


## Variable Summary

- measurement_date = date measurements were taken, including capture day
- collection/capture time for each lizard
- time_processed = when mass and blood draw were recorded
- time_c_temp = the time when cloacal temperature was recorded, immediately after CEWL measurements
- type = whether measurements were during experiment (exp) or after rehydration (post-rehab). For this R script/analysis, I'm only going to use capture day data, which is listed as "exp"
- day = whether measurements are from capture day or post-experiment, which was recorded in relation to CEWL & cloacal temp data. All observations used for this analysis will be from capture day
- individual ID for each lizard
- mass in grams
- hemolyzed = whether or not red blood cells burst and contaminated plasma
- hematocrit_percent = percent of blood that's red blood cells (measured in CRITOCAP microhematocrit capillary tubes)
- osmolality_mmol_kg_mean = the mean of 1-3 technical replicates of plasma osmolality measurements taken from plasma extracted from our blood samples and run on a VAPRO vapor pressure osmometer
- CEWL_g_m2h_mean = the mean of 3-5 technical replicates, after outliers were omitted, of CEWL measurements taken in the same area of the dorsum
- cloacal_temp_C = cloacal temperature recorded immediately after CEWL measurements
- capture_date = date of capture. For this dataset, it should be the same as measurement date
- day_n = numeric day of measurement. In this dataset, it should always be zero
- capture_date_time = combination of capture date and time
- SVL_mm = snout-to-vent length in mm



## Weather Data

This data was obtained from http://www.itrc.org/databases/precip/ (Adcon Server Data) to test the effect of ambient conditions on CEWL.

Load and format: 

```{r load weather data}
weather <- read.csv("./data/weather.csv", sep = ';') %>%
         # add a variable for combined date-time
  mutate(capture_date_time = as.POSIXct(paste(date, time),
                                format = "%m/%d/%y %I:%M %p"))
```

The weather data is only every 15 minutes, but I want to match it to any minute measurement, so I need to interpolate the values for each minute.

First, make a separate dataframe with every minute on each capture day.

```{r every minute data}
all_times <- data.frame(capture_date_time = c(
                           # June 16
                           seq(from = as.POSIXct("2021-06-16 07:00"),
                               to = as.POSIXct("2021-06-16 19:00"),
                               by="min"),
                           # June 26
                           seq(from = as.POSIXct("2021-06-26 07:00"),
                               to = as.POSIXct("2021-06-26 19:00"),
                               by="min"),
                           # July 20
                           seq(from = as.POSIXct("2021-07-20 07:00"),
                               to = as.POSIXct("2021-07-20 19:00"),
                               by="min"),
                           # August 8
                           seq(from = as.POSIXct("2021-08-08 07:00"),
                               to = as.POSIXct("2021-08-08 19:00"),
                               by="min"),
                           # August 22
                           seq(from = as.POSIXct("2021-08-22 07:00"),
                               to = as.POSIXct("2021-08-22 19:00"),
                               by="min")
                           ))

```

Next, merge the weather data into the times dataframe and interpolate the temperature and humidity between measurements.

```{r merge weather data}
weather_every_minute <- all_times %>% # time only dataframe
  # add weather measurements based on matching date-time
  left_join(weather, by = 'capture_date_time') %>%
         # convert temperature units F->C
  mutate(temp_C = fahrenheit.to.celsius(temperature_F, round = 2),
         # interpolate temperatures
         temp_C_interpol = na.approx(temp_C),
         # also get temperature C-> K
         temp_K_interpol = temp_C_interpol + 273.15,
         # interpolate humidities
         RH_percent_interpol = na.approx(relative_humidity_percent),
         # interpolate Wind Speeds
         wind_mph_interpol = na.approx(wind_speed_mph),
         # interpolate solar radiation
         solar_rad_W_sqm_interpol = na.approx(solar_radiation_W_sqm),
         # compute vapor pressure deficit
         # find saturation level first
         e_s_kPa_int = 0.611*exp((2500000/461.5)*
                                  ((1/273)-(1/temp_K_interpol))),
         # actual vapor pressure
         e_a_kPa_int = e_s_kPa_int * (RH_percent_interpol/100),
         # VPD
         VPD_kPa_int = e_s_kPa_int - e_a_kPa_int
         ) %>%
  # keep only the relevant variables
  dplyr::select(capture_date_time, 
                temp_C_interpol, 
                RH_percent_interpol, 
                VPD_kPa_int,
                wind_mph_interpol, 
                solar_rad_W_sqm_interpol)
summary(weather_every_minute)
```

I will add the weather data in when I add the scaled mass index (computed next) to the dataframe.



## Compute Scaled Mass Index

This is also known as the body condition index, or log-log residuals.

I calculate as described by: Peig, J., & Green, A. J. (2009). New perspectives for estimating body condition from mass/length data: The scaled mass index as an alternative method. Oikos, 118(12), 1883–1891. https://doi.org/10.1111/j.1600-0706.2009.17643.x

### Step 1: Simple Linear Regression

```{r SMI SLR}
mass_SVL_SLR <- lm(data = capture_dat, mass_g ~ SVL_mm)
summary(mass_SVL_SLR)
```


### Step 2: Identify Outliers

```{r SMI equation outliers 1}
plot(mass_SVL_SLR)
```

The conditions of linearity, equal error variance, and normality are all satisfied. It doesn't look like any residuals are >3 or <-3.

```{r SMI equation outliers 2}
boxplot(residuals(mass_SVL_SLR))
hist(residuals(mass_SVL_SLR))
```

From the boxplot, there is one individual with a much higher residual than the rest of the distribution. The histogram looks fine, and incredibly normally distributed.

Check average residual value:

```{r SMI equation outliers 3}
mean(residuals(mass_SVL_SLR))
median(residuals(mass_SVL_SLR))
```

The mean is basically zero and the median is pretty close to zero, which is very good.

Check for high leverage points:

```{r SMI equation outliers 4}
# compute values for observations 
high_leverage <- data.frame(H = hatvalues(mass_SVL_SLR)) %>% 
  mutate(row = row_number())

# compute cutoff value 
h_bar <- (3*sum(high_leverage$H))/nrow(high_leverage)

# add to original dataframe 
# see which observations have extremely high leverage (if any)
high_leverage_dat <- capture_dat %>%
  mutate(row = row_number()) %>%
  left_join(., high_leverage, by = "row") %>%
  dplyr::filter(H > h_bar) 
high_leverage_dat
```

No points are considered high leverage, which is fantastic.

Check for influential points based on Cook's distance:

```{r SMI equation outliers 5}
# get Cook's distance 
cooks <- data.frame(c = cooks.distance(mass_SVL_SLR)) %>% 
  mutate(row = row_number())

# add to original dataframe 
influential <- capture_dat %>%
  mutate(row = row_number()) %>% 
  left_join(., cooks, by = "row")

# see moderately influential points 
cook_mod_inf <- influential %>% 
  dplyr::filter(c>0.5) 
cook_mod_inf
```

There are no infuential points based on Cook's distance, so there's nothing to potentially remove.

We could remove the one outlier found using the boxplot, but it's the only one, so we will leave it in the dataset. No points were indicated to be outliers based on residuals or a histogram, and there were no high leverage or influential points. Thus I can create a log-log model using the data as-is. Observation omissions are unlikely to increase generalizability.


 
### Step 3: log-log Regression

```{r SMI log log regression}
log_mass_SVL_SLR <- lm(data = capture_dat, 
                       log(mass_g) ~ log(SVL_mm))
summary(log_mass_SVL_SLR)
```


### Step 4: Extract Values

compute standardized major axis using the log-log regression equation:

```{r SMI equation}
r <- sqrt(0.5699) # Pearson's correlection coefficient (sqrt of R-squared)
b_OLS <- 2.0013 # regression slope
b_SMA <- b_OLS/r
```

mean length in capture data:

```{r mean SVL}
L0 <- mean(capture_dat$SVL_mm)
```


### Step 5: Calculate Scaled Mass Index

(And join weather data.)

```{r join all data}
capture_dat_plus <- capture_dat %>%
  # compute SMI
  mutate(SMI = mass_g * ((L0/SVL_mm) ^ b_SMA)) %>%
  # join weather data
  left_join(weather_every_minute, by = c("capture_date_time")) %>%
  # clean up the dataframe
  dplyr::select(capture_date, capture_date_time, individual_ID, # basics
                mass_g, SVL_mm, SMI, # lizard size
                hemolyzed, hematocrit_percent, osmolality_mmol_kg_mean, # blood
                CEWL_g_m2h_mean, msmt_temp_C, msmt_VPD_kPa, cloacal_temp_C, # CEWL
                temp_C_interpol, VPD_kPa_int, wind_mph_interpol, solar_rad_W_sqm_interpol # weather
                )
summary(capture_dat_plus)
```


### Check

Look at the difference between regular mass and SMI:

```{r SMI vs mass plot}
capture_dat_plus %>% 
  ggplot(data = .) + 
  geom_point(aes(x = mass_g,
                 y = SMI, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = mass_g, 
                  y = SMI, 
                  ), 
              formula = y ~ x, 
              method = "lm", 
              color = "gray",
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Mass") + 
  ylab("SMI")
```






# Quick Plots

Plot very basic graphs to get an idea of what variables to incorporate into models and how.

## Osmolality

```{r basic osml plots}
plot(capture_dat_plus$individual_ID,
     capture_dat_plus$osmolality_mmol_kg_mean)
plot(capture_dat_plus$capture_date,
     capture_dat_plus$osmolality_mmol_kg_mean)
plot(capture_dat_plus$mass_g,
     capture_dat_plus$osmolality_mmol_kg_mean)
plot(capture_dat_plus$SVL_mm,
     capture_dat_plus$osmolality_mmol_kg_mean)
plot(capture_dat_plus$SMI,
     capture_dat_plus$osmolality_mmol_kg_mean)
plot(capture_dat_plus$hemolyzed,
     capture_dat_plus$osmolality_mmol_kg_mean)
plot(capture_dat_plus$hematocrit_percent,
     capture_dat_plus$osmolality_mmol_kg_mean)
plot(capture_dat_plus$CEWL_g_m2h_mean,
     capture_dat_plus$osmolality_mmol_kg_mean)
plot(capture_dat_plus$temp_C_interpol,
     capture_dat_plus$osmolality_mmol_kg_mean)
plot(capture_dat_plus$VPD_kPa_int,
     capture_dat_plus$osmolality_mmol_kg_mean)
plot(capture_dat_plus$wind_mph_interpol,
     capture_dat_plus$osmolality_mmol_kg_mean)
plot(capture_dat_plus$solar_rad_W_sqm_interpol,
     capture_dat_plus$osmolality_mmol_kg_mean)
```

There does not appear to be a meaningful visual trend for any predictors of plasma osmolality, so it will be interesting to see how the model selection process goes... There is definitely an increase in osmolality over the course of the season, though.


## CEWL

```{r basic CEWL plots}
plot(capture_dat_plus$individual_ID,
     capture_dat_plus$CEWL_g_m2h_mean)
plot(capture_dat_plus$capture_date,
     capture_dat_plus$CEWL_g_m2h_mean)
plot(capture_dat_plus$mass_g,
     capture_dat_plus$CEWL_g_m2h_mean)
plot(capture_dat_plus$SVL_mm,
     capture_dat_plus$CEWL_g_m2h_mean)
plot(capture_dat_plus$SMI,
     capture_dat_plus$CEWL_g_m2h_mean)
plot(capture_dat_plus$hematocrit_percent,
     capture_dat_plus$CEWL_g_m2h_mean)
plot(capture_dat_plus$osmolality_mmol_kg_mean,
     capture_dat_plus$CEWL_g_m2h_mean)
plot(capture_dat_plus$cloacal_temp_C,
     capture_dat_plus$CEWL_g_m2h_mean)
plot(capture_dat_plus$temp_C_interpol,
     capture_dat_plus$CEWL_g_m2h_mean)
plot(capture_dat_plus$VPD_kPa_int,
     capture_dat_plus$CEWL_g_m2h_mean)
plot(capture_dat_plus$wind_mph_interpol,
     capture_dat_plus$CEWL_g_m2h_mean)
plot(capture_dat_plus$solar_rad_W_sqm_interpol,
     capture_dat_plus$CEWL_g_m2h_mean)
```

It looks like there are meaningful differences in CEWL across individuals/dates (probably confounded), and based on cloacal temp, capture temp, capture VPD, capture wind, and capture solar radiation.



# LMMs


## Hematocrit


### Models

First, start with a full model with every probable potential predictor in it, then check for multicollinearity.

Hematocrit did not vary by capture day, so we are using simple linear models rather than linear mixed-effects models.

```{r hct model 1}
hct_mod1 <- lm(data = capture_dat_plus,
                          # response variable
                          hematocrit_percent ~ 
                          # body size
                          mass_g + SVL_mm + SMI +
                          # weather at the time of capture
                          temp_C_interpol * VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol
                            )
hct_mod1_VIFs <- data.frame(VIF = car::vif(hct_mod1)) %>%
  arrange(desc(VIF))
hct_mod1_VIFs
```

remove VPD*temp interaction:

```{r hct model 2}
hct_mod2 <- lm(data = capture_dat_plus,
                          # response variable
                          hematocrit_percent ~ 
                          # body size
                          mass_g + SVL_mm + SMI +
                          # weather at the time of capture
                          temp_C_interpol + VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol
                            )
hct_mod2_VIFs <- data.frame(VIF = car::vif(hct_mod2)) %>%
  arrange(desc(VIF))
hct_mod2_VIFs
drop1(hct_mod2)
```


drop SVL:

```{r hct model 3}
hct_mod3 <- lm(data = capture_dat_plus,
                          # response variable
                          hematocrit_percent ~ 
                          # body size
                          mass_g + SMI +
                          # weather at the time of capture
                          temp_C_interpol + VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol
                            )
hct_mod3_VIFs <- data.frame(VIF = car::vif(hct_mod3)) %>%
  arrange(desc(VIF))
hct_mod3_VIFs
drop1(hct_mod3)
```


drop temperature:

```{r hct model 4}
hct_mod4 <- lm(data = capture_dat_plus,
                          # response variable
                          hematocrit_percent ~ 
                          # body size
                          mass_g + SMI +
                          # weather at the time of capture
                          VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol
                            )
hct_mod4_VIFs <- data.frame(VIF = car::vif(hct_mod4)) %>%
  arrange(desc(VIF))
hct_mod4_VIFs
drop1(hct_mod4)
```

VIFs are all below 5 now, so start backwards selection.

Drop mass first:

```{r hct model 5}
hct_mod5 <- lm(data = capture_dat_plus,
                          # response variable
                          hematocrit_percent ~ 
                          # body size
                          SMI +
                          # weather at the time of capture
                          VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol
                            )
drop1(hct_mod5)
```


Drop VPD:

```{r hct model 6}
hct_mod6 <- lm(data = capture_dat_plus,
                          # response variable
                          hematocrit_percent ~ 
                          # body size
                          SMI +
                          # weather at the time of capture
                          wind_mph_interpol + solar_rad_W_sqm_interpol
                            )
drop1(hct_mod6)
```

Drop solar:

```{r hct model 7}
hct_mod7 <- lm(data = capture_dat_plus,
                          # response variable
                          hematocrit_percent ~ 
                          # body size
                          SMI +
                          # weather at the time of capture
                          wind_mph_interpol
                            )
drop1(hct_mod7)
```

Drop SMI:

```{r hct model 8}
hct_mod8 <- lm(data = capture_dat_plus,
                          # response variable
                          hematocrit_percent ~ 
                          # weather at the time of capture
                          wind_mph_interpol
                            )
```


Finally, null model:

```{r hct null model}
hct_mod_null <- lm(data = capture_dat_plus,
                          # response variable
                          hematocrit_percent ~ 1)
```


### Selection

Compare models 4-8 and the null model.

```{r hct mod compare}
hct_models <- list(hct_mod4, hct_mod5, hct_mod6, hct_mod7, 
                   hct_mod8, hct_mod_null)

#specify model names
hct_mod_names <- c('(model 4) ~ Wind-C, SMI, Solar-C, VPD-C, Mass', 
                       '(model 5) ~ Wind-C, SMI, Solar-C, VPD-C', 
                       '(model 6) ~ Wind-C, SMI, Solar-C',
                       '(model 7) ~ Wind-C, SMI', 
                       '(model 8) ~ Wind-C', 
                       'null model')
#calculate AIC of each model
hct_AICc <- data.frame(aictab(cand.set = hct_models, 
                                 modnames = hct_mod_names))
hct_AICc
```


The best models are models 6 and 5.


### LM Conditions

Check that the best model meets the criteria for linear regression and has no collinearity.

```{r hct model assumptions}
vif(hct_mod6)
plot(hct_mod6)
simple.eda(residuals(hct_mod6))
shapiro.test(residuals(hct_mod6))
vif(hct_mod5)
plot(hct_mod5)
simple.eda(residuals(hct_mod5))
shapiro.test(residuals(hct_mod5))
```

Everything is almost perfect.


### Export

```{r}
write.csv(hct_AICc, "./results_statistics/capture_hct_mod_rankings.csv")
write.csv(broom.mixed::tidy(hct_mod6), 
          "./results_statistics/capture_hct_best_mod1.csv")
write.csv(broom.mixed::tidy(hct_mod5), 
          "./results_statistics/capture_hct_best_mod2.csv")
```


## Osmolality


### Models

Since there are large differences in osmolality by date, but we are interested in what's different among dates, rather than the capture date itself, we will include that as a random effect in the model.

We would also include whether or not a blood sample is hemolyzed as a random effect, but only 11 of the almost 150 samples were hemolyzed, so we will assume that any potential effects will be undetectable and/or overshadowed. We do not have concern about using those points.

First, start with a full model with every probable potential predictor in it, then check for multicollinearity.

```{r osmolality model 1}
osml_mod1 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          osmolality_mmol_kg_mean ~ 
                          # body size
                          mass_g + SVL_mm + SMI +
                          # blood sample traits
                          hematocrit_percent + 
                          # weather at the time of capture
                          temp_C_interpol * VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol +
                          (1|capture_date))
osml_mod1_VIFs <- data.frame(VIF = car::vif(osml_mod1)) %>%
  arrange(desc(VIF))
osml_mod1_VIFs
```

VPD and temperature introduce a lot of collinearity, so start by dropping their interaction:

```{r osmolality model 2}
osml_mod2 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          osmolality_mmol_kg_mean ~ 
                          # body size
                          mass_g + SVL_mm + SMI +
                          # blood sample traits
                          hematocrit_percent + 
                          # weather at the time of capture
                          temp_C_interpol + VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol +
                          (1|capture_date))
osml_mod2_VIFs <- data.frame(VIF = car::vif(osml_mod2)) %>%
  arrange(desc(VIF))
osml_mod2_VIFs
drop1(osml_mod2)
```

Drop mass next, since it's extremely collinear and we get slightly better AIC by dropping mass compared to SVL:

```{r osmolality model 3}
osml_mod3 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          osmolality_mmol_kg_mean ~ 
                          # body size
                          SVL_mm + SMI +
                          # blood sample traits
                          hematocrit_percent + 
                          # weather at the time of capture
                          temp_C_interpol + VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol +
                          (1|capture_date))
osml_mod3_VIFs <- data.frame(VIF = car::vif(osml_mod3)) %>%
  arrange(desc(VIF))
osml_mod3_VIFs
drop1(osml_mod3)
```

Temperature is still introducing a lot of multicollinearity, so drop:

```{r osmolality model 4}
osml_mod4 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          osmolality_mmol_kg_mean ~ 
                          # body size
                          SVL_mm + SMI +
                          # blood sample traits
                          hematocrit_percent + 
                          # weather at the time of capture
                          VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol +
                          (1|capture_date))
osml_mod4_VIFs <- data.frame(VIF = car::vif(osml_mod4)) %>%
  arrange(desc(VIF))
osml_mod4_VIFs
summary(osml_mod4)
drop1(osml_mod4)
```


Great, VIFs are well-within acceptable ranges. Now we can start backwards model selection.

Start by dropping wind:

```{r osmolality model 5}
osml_mod5 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          osmolality_mmol_kg_mean ~ 
                          # body size
                          SVL_mm + SMI +
                          # blood sample traits
                          hematocrit_percent + 
                          # weather at the time of capture
                          VPD_kPa_int + solar_rad_W_sqm_interpol +
                          (1|capture_date))
summary(osml_mod5)
drop1(osml_mod5)
```

Drop hematocrit:

```{r osmolality model 6}
osml_mod6 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          osmolality_mmol_kg_mean ~ 
                          # body size
                          SVL_mm + SMI +
                          # weather at the time of capture
                          VPD_kPa_int + solar_rad_W_sqm_interpol +
                          (1|capture_date))
summary(osml_mod6)
drop1(osml_mod6)
```

Drop SMI:

```{r osmolality model 7}
osml_mod7 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          osmolality_mmol_kg_mean ~ 
                          # body size
                          SVL_mm +
                          # weather at the time of capture
                          VPD_kPa_int + solar_rad_W_sqm_interpol +
                          (1|capture_date))
summary(osml_mod7)
drop1(osml_mod7)
```

Drop VPD:

```{r osmolality model 8}
osml_mod8 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          osmolality_mmol_kg_mean ~ 
                          # body size
                          SVL_mm +
                          # weather at the time of capture
                          solar_rad_W_sqm_interpol +
                          (1|capture_date))
summary(osml_mod8)
```

Drop SVL:

```{r osmolality model 9}
osml_mod9 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          osmolality_mmol_kg_mean ~ 
                          # weather at the time of capture
                          solar_rad_W_sqm_interpol +
                          (1|capture_date))
summary(osml_mod9)
```

Lastly, compute null model:

```{r osmolality null model}
osml_mod_null <- lme4::lmer(data = capture_dat_plus,
                          osmolality_mmol_kg_mean ~ 1 +
                          (1|capture_date))
summary(osml_mod_null)
```


### Selection

Compare models 4-9 and null. 

```{r osml mod compare}
osml_models <- list(osml_mod4, osml_mod5, osml_mod6,
                    osml_mod7, osml_mod8, osml_mod9,
                    osml_mod_null)
#specify model names
osml_mod_names <- c('(model 4) ~ Solar-C, SVL, VPD-C, SMI, Hct, Wind-C', 
                       '(model 5) ~ Solar-C, SVL, VPD-C, SMI, Hct', 
                       '(model 6) ~ Solar-C, SVL, VPD-C, SMI',
                       '(model 7) ~ Solar-C, SVL, VPD-C', 
                       '(model 8) ~ Solar-C, SVL', 
                       '(model 9) ~ Solar-C', 
                       'null model')
#calculate AIC of each model
osml_AICc <- data.frame(aictab(cand.set = osml_models, 
                                 modnames = osml_mod_names))
osml_AICc
```




### LM Conditions

Check residual plots and VIFs

```{r osml model assumptions}
vif(osml_mod6)
plot(osml_mod6)
simple.eda(residuals(osml_mod6))
shapiro.test(residuals(osml_mod6))
vif(osml_mod7)
plot(osml_mod7)
simple.eda(residuals(osml_mod7))
shapiro.test(residuals(osml_mod7))
```

There is no clear pattern in the residuals ~ fitted plot, so linearity seems satisfied. slight fanning, but equal error variance seems fine. Normality seems fine, even though the Shapiro-Wilk normality test is significant. VIFs essentially negligible.


### Export


First, re-run for p-values:


```{r osmolality best models}
osml_mod6p <- lmerTest::lmer(data = capture_dat_plus,
                          # response variable
                          osmolality_mmol_kg_mean ~ 
                          # body size
                          SVL_mm + SMI +
                          # weather at the time of capture
                          VPD_kPa_int + solar_rad_W_sqm_interpol +
                          (1|capture_date))
summary(osml_mod6p)

osml_mod7p <- lmerTest::lmer(data = capture_dat_plus,
                          # response variable
                          osmolality_mmol_kg_mean ~ 
                          # body size
                          SVL_mm +
                          # weather at the time of capture
                          VPD_kPa_int + solar_rad_W_sqm_interpol +
                          (1|capture_date))
summary(osml_mod7p)
```


Save the model output.

```{r best osml model}
write.csv(broom.mixed::tidy(osml_mod6p),
          "./results_statistics/capture_osml_best_model1.csv")
write.csv(broom.mixed::tidy(osml_mod7p),
          "./results_statistics/capture_osml_best_model2.csv")
write.csv(osml_AICc, "./results_statistics/capture_osml_mod_rankings.csv")
```


To report in paper:

The best models to predict the variation in baseline plasma osmolality included SVL, SMI, VPD, and solar radiation at the time of capture as fixed effects. Date was included as a random effect. The final model had acceptable LM conditions. The full model included mass, SVL, SMI, percent hematocrit, and temperature, VPD, wind speed, and solar radiation at the time of capture, with date as a random effect.



## CEWL

It looks like there are meaningful differences in CEWL across individuals/dates (probably confounded), and based on cloacal temp, capture temp, capture VPD, capture wind, and capture solar radiation.

*I didn't calculate holding time*!!!!!!!!!!!!!!!!!!!

### Models

Start with the full model of all potential predictor variables. We will again include date as a random effect. Individual ID is not included as a random effect bc each lizard only has one set of capture observations.

When we have this many variables, it's extremely important to start with checking for multicollinearity.

```{r CEWL model 1}
CEWL_mod1 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # essential covariate
                          cloacal_temp_C +
                          # body size
                          mass_g + SVL_mm + SMI +
                          # blood
                          osmolality_mmol_kg_mean + hematocrit_percent + 
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          # weather at the time of capture
                          temp_C_interpol * VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol +
                          (1|capture_date))
CEWL_mod1_VIFs <- data.frame(VIF = car::vif(CEWL_mod1)) %>%
  arrange(desc(VIF))
CEWL_mod1_VIFs
drop1(CEWL_mod1)
```


Just as for osmolality, VPD and temperature introduce a lot of collinearity. Start with dropping their interaction:

```{r CEWL model 2}
CEWL_mod2 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # essential covariate
                          cloacal_temp_C +
                          # body size
                          mass_g + SVL_mm + SMI +
                          # blood
                          osmolality_mmol_kg_mean + hematocrit_percent + 
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          # weather at the time of capture
                          temp_C_interpol + VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol +
                          (1|capture_date))
CEWL_mod2_VIFs <- data.frame(VIF = car::vif(CEWL_mod2)) %>%
  arrange(desc(VIF))
CEWL_mod2_VIFs
drop1(CEWL_mod2)
```

MUCH better. Drop SVL next:

```{r CEWL model 3}
CEWL_mod3 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # essential covariate
                          cloacal_temp_C +
                          # body size
                          mass_g + SMI +
                          # blood
                          osmolality_mmol_kg_mean + hematocrit_percent + 
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          # weather at the time of capture
                          temp_C_interpol + VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol +
                          (1|capture_date))
CEWL_mod3_VIFs <- data.frame(VIF = car::vif(CEWL_mod3)) %>%
  arrange(desc(VIF))
CEWL_mod3_VIFs
drop1(CEWL_mod3)
```


Next drop temperature at the time of capture:

```{r CEWL model 4}
CEWL_mod4 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # essential covariate
                          cloacal_temp_C +
                          # body size
                          mass_g + SMI +
                          # blood
                          osmolality_mmol_kg_mean + hematocrit_percent + 
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          # weather at the time of capture
                          VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol +
                          (1|capture_date))
CEWL_mod4_VIFs <- data.frame(VIF = car::vif(CEWL_mod4)) %>%
  arrange(desc(VIF))
CEWL_mod4_VIFs
summary(CEWL_mod4)
drop1(CEWL_mod4)
```


Great, VIFs are minimal and we're ready to start backwards selection!

Start with dropping SMI:

```{r CEWL model 5}
CEWL_mod5 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # essential covariate
                          cloacal_temp_C +
                          # body size
                          mass_g + 
                          # blood
                          osmolality_mmol_kg_mean + hematocrit_percent + 
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          # weather at the time of capture
                          VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol +
                          (1|capture_date))
summary(CEWL_mod5)
drop1(CEWL_mod5)
```

next drop cloacal temperature (What?!):

```{r CEWL model 6}
CEWL_mod6 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # body size
                          mass_g + 
                          # blood
                          osmolality_mmol_kg_mean + hematocrit_percent + 
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          # weather at the time of capture
                          VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol +
                          (1|capture_date))
summary(CEWL_mod6)
drop1(CEWL_mod6)
```


next drop hematocrit:

```{r CEWL model 7}
CEWL_mod7 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # body size
                          mass_g + 
                          # blood
                          osmolality_mmol_kg_mean +
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          # weather at the time of capture
                          VPD_kPa_int +
                          wind_mph_interpol + solar_rad_W_sqm_interpol +
                          (1|capture_date))
summary(CEWL_mod7)
drop1(CEWL_mod7)
```


next drop solar radiation:

```{r CEWL model 8}
CEWL_mod8 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # body size
                          mass_g + 
                          # blood
                          osmolality_mmol_kg_mean +
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          # weather at the time of capture
                          VPD_kPa_int +
                          wind_mph_interpol +
                          (1|capture_date))
summary(CEWL_mod8)
drop1(CEWL_mod8)
```


next drop mass:

```{r CEWL model 9}
CEWL_mod9 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # blood
                          osmolality_mmol_kg_mean +
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          # weather at the time of capture
                          VPD_kPa_int +
                          wind_mph_interpol +
                          (1|capture_date))
summary(CEWL_mod9)
drop1(CEWL_mod9)
```

next drop VPD:

```{r CEWL model 10}
CEWL_mod10 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # blood
                          osmolality_mmol_kg_mean +
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          # weather at the time of capture
                          wind_mph_interpol +
                          (1|capture_date))
summary(CEWL_mod10)
# can't test drop1 bc of NA's 
```

based on t-values, wind should be dropped next:

```{r CEWL model 11}
CEWL_mod11 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # blood
                          osmolality_mmol_kg_mean +
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          (1|capture_date))
summary(CEWL_mod11)
drop1(CEWL_mod11)
```

drop osmolality:

```{r CEWL model 12}
CEWL_mod12 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          (1|capture_date))
summary(CEWL_mod12)
drop1(CEWL_mod12)
```


drop VPD at the time of msmt:

```{r CEWL model 13}
CEWL_mod13 <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # microclimate at the time of msmt
                          msmt_temp_C +
                          (1|capture_date))
summary(CEWL_mod13)
```

And finally, null model:

```{r CEWL model null}
CEWL_mod_null <- lme4::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 1 +
                          (1|capture_date))
summary(CEWL_mod_null)
```



### Selection

compare models 4-13 and the null

```{r CEWL mod compare}
CEWL_models <- list(CEWL_mod4, CEWL_mod5, CEWL_mod6, CEWL_mod7, 
                    CEWL_mod8, CEWL_mod9, CEWL_mod10, CEWL_mod11,
                    CEWL_mod12, CEWL_mod13, CEWL_mod_null)

#specify model names
CEWL_mod_names <- c('(model 4) ~ Temp-M, VPD-M, Osml, Wind-C, VPD-C, Mass, Solar-C, Hct, Temp-Clo, SMI', 
                    '(model 5) ~ Temp-M, VPD-M, Osml, Wind-C, VPD-C, Mass, Solar-C, Hct, Temp-Clo', 
                    '(model 6) ~ Temp-M, VPD-M, Osml, Wind-C, VPD-C, Mass, Solar-C, Hct', 
                       '(model 7) ~ Temp-M, VPD-M, Osml, Wind-C, VPD-C, Mass, Solar-C', 
                       '(model 8) ~ Temp-M, VPD-M, Osml, Wind-C, VPD-C, Mass',
                       '(model 9) ~ Temp-M, VPD-M, Osml, Wind-C, VPD-C', 
                       '(model 10) ~ Temp-M, VPD-M, Osml, Wind-C', 
                       '(model 11) ~ Temp-M, VPD-M, Osml',
                      '(model 12) ~ Temp-M, VPD-M',
                      '(model 13) ~ Temp-M',
                      'null model')
#calculate AIC of each model
CEWL_AICc <- data.frame(aictab(cand.set = CEWL_models, 
                                 modnames = CEWL_mod_names))
CEWL_AICc
```


The single best model is model 9, which included Temp-M, VPD-M, Osml, Wind-C, VPD-C.



### LM Conditions

Check that the best model meets the criteria for linear regression and has no collinearity.

```{r CEWL model assumptions}
vif(CEWL_mod9)
plot(CEWL_mod9)
simple.eda(residuals(CEWL_mod9))
shapiro.test(residuals(CEWL_mod9))
```

There is some slight fanning in the residuals ~ fitted plot, suggesting equal error variance is not perfect, but overall, all LNE conditions appear to be met and VIFs are very low.


### Export

First, re-run the best model using lmerTest for p-values.

```{r CEWL model 9 p}
CEWL_mod9p <- lmerTest::lmer(data = capture_dat_plus,
                          # response variable
                          CEWL_g_m2h_mean ~ 
                          # blood
                          osmolality_mmol_kg_mean +
                          # microclimate at the time of msmt
                          msmt_temp_C + msmt_VPD_kPa +
                          # weather at the time of capture
                          VPD_kPa_int +
                          wind_mph_interpol +
                          (1|capture_date))
summary(CEWL_mod9p)
```

Save the best CEWL model output.

```{r best CEWL model}
write.csv(broom.mixed::tidy(CEWL_mod9p),
          "./results_statistics/capture_CEWL_best_model.csv")
write.csv(CEWL_AICc, 
          "./results_statistics/capture_CEWL_mod_rankings.csv")
```


To report in paper:

The best model to predict CEWL included plasma osmolality, temperature and VPD at the time of measurement, and VPD and wind at the time of capture. The final model met all linear regression conditions for linearity, normality, and equal error variance, and there was no multicollinearity.


# Pub Figures

## Custom Colors

```{r}
lizard_color = "turquoise"
VPD_color = "blue"
temp_color = "gray"
solar_color = "orange"
wind_color = "orange"
date_color = "gray"
osml_color = "red"
```



## Hct ~ SMI

```{r hct SMI fig}
ggplot(capture_dat_plus) + 
  aes(x = SMI,
      y = hematocrit_percent) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              color = lizard_color,
              alpha = 1) + 
  theme_classic() + 
  xlab("Body Condition (g)") + 
  ylab("Hematocrit (%)") + 
  #ylab("") +
  #xlim() +
  ylim(20,60) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_hct_SMI_fig
cap_hct_SMI_fig
```



## Hct ~ VPD at Capture

```{r hct VPD fig}
ggplot(capture_dat_plus) + 
  aes(x = VPD_kPa_int,
      y = hematocrit_percent) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = VPD_color,
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("VPD at Capture (kPa)") + 
  ylab("Hematocrit (%)") + 
  #ylab("") +
  #xlim() +
  ylim(20,60) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_hct_VPD_fig
cap_hct_VPD_fig
```



## Hct ~ Wind Speed at Capture

```{r hct wind fig}
ggplot(capture_dat_plus) + 
  aes(x = wind_mph_interpol,
      y = hematocrit_percent)+
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = wind_color,
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Wind Speed at Capture (mph)") + 
  ylab("Hematocrit (%)") + 
  #ylab("") +
  #xlim() +
  ylim(20, 60) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_hct_wind_fig
cap_hct_wind_fig
```




## Hct ~ Solar Radiation at Capture

```{r hct sorad fig}
ggplot(capture_dat_plus) + 
  aes(x = solar_rad_W_sqm_interpol,
      y = hematocrit_percent) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = solar_color,
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab(bquote('Solar Radiation (W/'*m^2*')')) + 
  ylab("Hematocrit (%)") + 
  #ylab("") +
  ylim(20, 60) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_hct_sorad_fig
cap_hct_sorad_fig
```


## Osmolality ~ SVL

```{r osml SVL fig}
ggplot(capture_dat_plus) + 
  aes(x = SVL_mm,
      y = osmolality_mmol_kg_mean) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              color = lizard_color,
              alpha = 1) + 
  theme_classic() + 
  xlab("Snout-Vent Length (mm)") + 
  ylab("Plasma Osmolality (mmol/kg)") + 
  #ylab("") +
  #xlim() +
  ylim(300,400) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_osml_SVL_fig
cap_osml_SVL_fig
```


## Osmolality ~ SMI

```{r osml SMI fig}
ggplot(capture_dat_plus) + 
  aes(x = SMI,
      y = osmolality_mmol_kg_mean) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              color = lizard_color,
              alpha = 1) + 
  theme_classic() + 
  xlab("Body Condition (g)") + 
  ylab("Plasma Osmolality (mmol/kg)") + 
  #ylab("") +
  #xlim() +
  ylim(300,400) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_osml_SMI_fig
cap_osml_SMI_fig
```


## Osmolality ~ VPD at Capture

```{r osml VPD fig}
ggplot(capture_dat_plus) + 
  aes(x = VPD_kPa_int,
      y = osmolality_mmol_kg_mean) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = VPD_color,
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("VPD at Capture (kPa)") + 
  ylab("Plasma Osmolality (mmol/kg)") + 
  #ylab("") +
  #xlim() +
  ylim(300,400) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_osml_VPD_fig
cap_osml_VPD_fig
```


## Osmolality ~ Solar Radiation at Capture

```{r osml sorad fig}
ggplot(capture_dat_plus) + 
  aes(x = solar_rad_W_sqm_interpol,
      y = osmolality_mmol_kg_mean) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = solar_color,
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab(bquote('Solar Radiation (W/'*m^2*')')) + 
  ylab("Plasma Osmolality (mmol/kg)") + 
  #ylab("") +
  ylim(300,400) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_osml_sorad_fig
cap_osml_sorad_fig
```


## Osmolality ~ Date

```{r osml date fig}
ggplot(capture_dat_plus) + 
  aes(x = as.factor(capture_date),
      y = osmolality_mmol_kg_mean,
      group = as.factor(capture_date)) +
  geom_boxplot() +
  geom_jitter(size = 1, 
              alpha = 0.4) + 
  theme_classic() + 
  xlab("Date") + 
  ylab("Plasma Osmolality (mmol/kg)") + 
  #ylab("") +
  #xlim() +
  ylim(300, 400) +
  #annotate("text", x = , y = , 
   #        label = "paste(italic(R) ^ 2, \" = 0.\")", 
    #       parse = TRUE,
     #      size = 6) +
  #annotate("text", x = , y = , 
   #        label = "paste(italic(p), \" < 0.0001\")", 
    #       parse = TRUE,
     #      size = 6) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_osml_date_fig
cap_osml_date_fig
```









## CEWL ~ Cloacal Temperature

```{r CEWL cloacal temp fig}
ggplot(capture_dat_plus) + 
  aes(x = cloacal_temp_C,
      y = CEWL_g_m2h_mean) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = temp_color,
              size = 1.6, 
              alpha = 1) + 
  theme_classic() + 
  xlab("Cloacal Temperature (°C)") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  #ylab("") +
  #xlim() +
  ylim(0, 40) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_CEWL_clotemp_fig
cap_CEWL_clotemp_fig
```


## CEWL ~ Plasma Osmolality

```{r CEWL osml fig}
ggplot(capture_dat_plus) + 
  aes(x = osmolality_mmol_kg_mean,
                 y = CEWL_g_m2h_mean) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = osml_color,
              size = 1.6, 
              alpha = 1) + 
  theme_classic() + 
  xlab("Plasma Osmolality (mmol/kg)") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  #ylab("") +
  xlim(300, 400) +
  ylim(0, 40) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_CEWL_osml_fig
cap_CEWL_osml_fig
```



## CEWL ~ Temperature at Measurement

```{r CEWL temp fig}
ggplot(capture_dat_plus) + 
  aes(x = msmt_temp_C,
      y = CEWL_g_m2h_mean) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = temp_color,
              size = 1.6, 
              alpha = 1) + 
  theme_classic() + 
  xlab("Temperature at Measurement (°C)") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  #ylab("") +
  #xlim() +
  ylim(0, 40) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_CEWL_temp_fig
cap_CEWL_temp_fig
```


## CEWL ~ VPD at Measurement

```{r CEWL VPD-m fig}
ggplot(capture_dat_plus) + 
  aes(x = msmt_VPD_kPa,
      y = CEWL_g_m2h_mean) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = VPD_color,
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("VPD at Measurement (kPa)") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  #ylab("") +
  #xlim() +
  ylim(0, 40) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_CEWL_VPDm_fig
cap_CEWL_VPDm_fig
```




## CEWL ~ VPD at Capture

```{r CEWL VPD-c fig}
ggplot(capture_dat_plus) + 
  aes(x = VPD_kPa_int,
      y = CEWL_g_m2h_mean) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = VPD_color,
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("VPD at Capture (kPa)") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  #ylab("") +
  #xlim() +
  ylim(0, 40) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_CEWL_VPDc_fig
cap_CEWL_VPDc_fig
```


## CEWL ~ Wind at Capture

```{r CEWL wind fig}
ggplot(capture_dat_plus) + 
  aes(x = wind_mph_interpol,
      y = CEWL_g_m2h_mean)+
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = wind_color,
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Wind Speed at Capture (mph)") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  #ylab("") +
  #xlim() +
  ylim(0, 40) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_CEWL_wind_fig
cap_CEWL_wind_fig
```


## CEWL ~ Date

```{r CEWL date fig}
ggplot(capture_dat_plus) + 
  aes(x = as.factor(capture_date),
      y = CEWL_g_m2h_mean,
      group = as.factor(capture_date)) +
  geom_boxplot() +
  geom_jitter(size = 1, 
              alpha = 0.4) + 
  theme_classic() + 
  xlab("Date") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  #ylab("") +
  #xlim() +
  ylim(0, 40) +
  #annotate("text", x = , y = , 
   #        label = "paste(italic(R) ^ 2, \" = 0.\")", 
    #       parse = TRUE,
     #      size = 6) +
  #annotate("text", x = , y = , 
   #        label = "paste(italic(p), \" < 0.0001\")", 
    #       parse = TRUE,
     #      size = 6) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        #axis.text.y = element_blank(),
        #plot.margin = unit(c(0.1,0,0.1,0.45), "cm")
        ) -> cap_CEWL_date_fig
cap_CEWL_date_fig
```




## Figure Arrangements

```{r FIG ARRANGE}
# hematocrit
ggarrange(cap_hct_sorad_fig, cap_hct_VPD_fig,
          cap_hct_SMI_fig, cap_hct_wind_fig,
          ncol = 2, nrow = 2,
          legend = "none"
          ) -> cap_hct_multi_fig
cap_hct_multi_fig
# export figure
ggsave(filename = "cap_hct_multi_fig.jpeg",
       plot = cap_hct_multi_fig,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 12, height = 8)


# osmolality
ggarrange(cap_osml_sorad_fig, cap_osml_VPD_fig,
          cap_osml_SMI_fig, cap_osml_SVL_fig,
          ncol = 2, nrow = 2,
          legend = "none"
          ) -> cap_osml_multi_fig
cap_osml_multi_fig
# export figure
ggsave(filename = "cap_osml_multi_fig.jpeg",
       plot = cap_osml_multi_fig,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 12, height = 8)



# CEWL
ggarrange(cap_CEWL_VPDm_fig, cap_CEWL_VPDc_fig,
          cap_CEWL_temp_fig, cap_CEWL_wind_fig,
          cap_CEWL_osml_fig, cap_CEWL_clotemp_fig,
          ncol = 2, nrow = 3,
          legend = "none"
          ) -> cap_CEWL_multi_fig
cap_CEWL_multi_fig
# export figure
ggsave(filename = "cap_CEWL_multi_fig",
       plot = cap_CEWL_multi_fig,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 12, height = 16)



# date differences
ggarrange(cap_osml_date_fig, cap_CEWL_date_fig,
          ncol = 1, nrow = 2,
          legend = "none"
          ) -> cap_date_diffs_multi_fig
cap_date_diffs_multi_fig
# export figure
ggsave(filename = "cap_date_diffs_multi_fig.jpeg",
       plot = cap_date_diffs_multi_fig,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 6, height = 8)
```















