---
title: "Climate Water Loss Experiment - HOBO Data Logger Climate Summary"
author: "Savannah Weaver"
date: "2022"
output: 
  rmdformats::html_clean:
    highlight: tango
    thumbnails: FALSE
    toc: TRUE
    toc_depth: 3
---

# Packages

```{r setup, include = TRUE, message = FALSE}
if (!require("tidyverse")) install.packages("tidyverse") 
library("tidyverse")
if (!require("rmdformats")) install.packages("rmdformats")
library("rmdformats") # clean html R markdown format
```


# Background

This data was collected using Onset HOBO temperature and humidity dataloggers to assess the climates that lizards actually experienced during the course of our acclimation experiment. 


# Load in Data


## File Names

The data is in a separate file for each download for each logger, so I need to compile each of those into one dataset. To do this, first I compile a list of the filenames I need to read-in.


```{r, filenames}
# make a list of file names of all data to load in
filenames <- list.files(path = "data/HOBOs")
```




## Function

Next, I make a function that will read in the data from each csv, name and organize the data correctly. 

```{r, create read_HOBO_files function to get each csv data}
# make a function to read in data from each csv file and add correct identifiers
read_HOBO_files <- function(filename) {
  
  # edit the filename inputted to funtion
  # to make a unique identifier for each logger
  name <- substr(filename, 1, nchar(filename)-15)
  # read in the csv file for this given filename
  dat <- read.csv(file.path("data/HOBOs", filename),
                # each csv has headers
                header = TRUE,
                # this is what I want to rename the col headers
                col.names = c("order", "date_time_PST", "temp_C", 
                              "relative_humidity", "dew_pt_C", 
                              # the 6,7,8th cols are not data
                              # logger use info we don't need
                              "mostly_blank", "mostly_blank", "mostly_blank")
                ) %>%
    # select only the cols with data we want
    # don't need order- just an arbitrary observation identifier
    # don't need "mostly_blank" cols- unnecessary logger use info
    # but get the rest of the cols with informative data
    dplyr::select(date_time_PST, temp_C, relative_humidity, dew_pt_C) %>%
    # add a column with the name of the HOBO the data is from 
    dplyr::mutate(HOBO_ID = name)
  
  # return the dataframe for that single csv file
  dat
}
```



## Apply

Finally, I apply the function I made to all of the filenames I compiled, then put all of those dataframes into one dataframe for my analyses.

This will print warnings saying that header and col.names are different lengths, because the data has extra notes on logger usage that we read-in, but get rid of.

```{r, apply read_HOBO_files function}
# apply function to get data from all csvs
all_HOBO_data <- lapply(filenames, read_HOBO_files) %>%
  # paste all data files together into one df by row
  reduce(rbind) %>% 
  mutate(date_time_PST = as.POSIXct(date_time_PST, 
                                    format = "%m/%d/%Y %H:%M:%S")) %>%
  # remove missing data
  dplyr::filter(complete.cases(date_time_PST, relative_humidity)) %>%
  mutate(serial = (substr(HOBO_ID, 7, 15)), # select characters
         serial = str_trim(serial), # remove trailing white space
         serial = str_replace_all(serial, "[^A-z0-9]", "_"), # replace any special characters with underscores
         serial = as.factor(serial), # set class as factor
         temp_C = as.numeric(temp_C),
         relative_humidity = as.numeric(relative_humidity),
         dew_pt_C = as.numeric(dew_pt_C)
         )
summary(all_HOBO_data)
head(all_HOBO_data)
```



# Data Wrangling


## Dates

The loggers were left on sometimes even though the experiment was not ongoing. I need to subset the data based on the date range for actual experiment days.

Trail 1: June 16-24
Trail 2: June 26 - July 4
Trial 3: July 20-28
Trial 4: August 8-16
Trial 5: August 22-30

Lizards were usually placed in the chambers late on the first day and taken out early on the last day, so I'll only use from 8 pm on the first day up to 6 am the last day.

```{r}
t1_days <- seq(as.POSIXct("2021-06-16 20:00:00"),
               as.POSIXct("2021-06-24 06:00:00"), "1 min")
t2_days <- seq(as.POSIXct("2021-06-26 20:00:00"),
               as.POSIXct("2021-07-04 06:00:00"), "1 min")
t3_days <- seq(as.POSIXct("2021-07-20 20:00:00"),
               as.POSIXct("2021-07-28 06:00:00"), "1 min")
t4_days <- seq(as.POSIXct("2021-08-08 20:00:00"),
               as.POSIXct("2021-08-16 06:00:00"), "1 min")
t5_days <- seq(as.POSIXct("2021-08-22 20:00:00"),
               as.POSIXct("2021-08-30 06:00:00"), "1 min")
exp_dates <- c(t1_days, t2_days, t3_days, t4_days, t5_days)
```

Subset data:

```{r}
subset_HOBO_data <- all_HOBO_data %>%
  dplyr::filter(date_time_PST %in% exp_dates)
```


## Trial & Tmt

Add our notes about what bins/treatments the loggers were put in.

```{r}
# load HOBO assignment data
HOBO_tmts <- read.csv("./data/HOBO_assignments.csv") %>%
  dplyr::filter(trial != "5 rehab") %>%
  mutate(trial = as.factor(trial),
         serial = as.factor(serial),
         humidity_tmt = as.factor(humidity_tmt),
         temp_tmt = as.factor(temp_tmt),
         tmt = as.factor(paste(temp_tmt, humidity_tmt, sep = "-"))
         ) %>%
  dplyr::select(-notes, -in.google.drive.) 
# combine
format_HOBO_data <- subset_HOBO_data %>% 
  mutate(trial = as.factor(case_when(date_time_PST %in% t1_days ~ 1,
                                     date_time_PST %in% t2_days ~ 2,
                                     date_time_PST %in% t3_days ~ 3,
                                     date_time_PST %in% t4_days ~ 4,
                                     date_time_PST %in% t5_days ~ 5)),
         date_only = as.Date(date_time_PST)
         ) %>%
  left_join(HOBO_tmts, by = c("trial", "serial"))
summary(format_HOBO_data)
```



# Check Data

## Dates

```{r}
format_HOBO_data %>% 
  group_by(trial) %>%
  summarise(min_date = (min(date_time_PST)),
            max_date = (max(date_time_PST)))
```


## Temperature

```{r}
hist(format_HOBO_data$temp_C)
format_HOBO_data %>%
  group_by(trial, serial, tmt) %>%
  summarise(max(temp_C),
            mean(temp_C),
            min(temp_C)
            ) %>%
  arrange(tmt)
```



```{r}
format_HOBO_data %>%
  ggplot() +
  geom_line(aes(x = date_time_PST,
                y = temp_C,
                color = tmt)) + 
  theme_classic() + 
  geom_hline(yintercept = 30) + 
  geom_hline(yintercept = 20)
```

There's a chance the values <30 for the hot tmt and <20 for the cool tmt are erroneous values.

```{r}
format_HOBO_data %>%
  dplyr::filter(temp_tmt == "hot" & temp_C < 30) %>%
  arrange(temp_C)
format_HOBO_data %>%
  dplyr::filter(temp_tmt == "cool" & temp_C < 20) %>%
  arrange(temp_C)
```

There are only 10 points <30C for the hot treatment, which is very few compared to the total number of measurements. I think these observations are from when the lizards (thus loggers) were out of the incubators for daily checkups. But, this is still technically characteristic of the environment the lizards experienced, and it's very few points that are unlikely to affect the summary statistics, so I'll leave them.

There are many more points <20C for the cool treatment. The lowest is still ~17C, and most of them are very close to 20. The bottom of the cool incubator tended to get extra cold, so I'm not worried about the validity of these points.

All the temperature observations are valid.



## Humidity

```{r}
hist(format_HOBO_data$relative_humidity)
format_HOBO_data %>%
  group_by(trial, serial, tmt) %>%
  summarise(max(relative_humidity),
            mean(relative_humidity),
            min(relative_humidity)
            ) %>%
  arrange(tmt)
```



```{r}
format_HOBO_data %>%
  ggplot() +
  geom_line(aes(x = date_time_PST,
                y = relative_humidity,
                color = tmt)) + 
  theme_classic()
```





# Statistics


## Treatment Differences

Get mean and sd of temperature and humidity for each tmt group, and calculate vapor pressure deficit based on those mean values using the formula in the book "Environmental Biophysics" by Campbell & Norman, 1998.

```{r}
tmt_only_means <- format_HOBO_data %>%
  group_by(tmt) %>%
  summarise(temp_mean = mean(temp_C),
            temp_SD = sd(temp_C),
            humidity_mean = mean(relative_humidity),
            humidity_SD = sd(relative_humidity)) %>%
  mutate(e_s_kPa = 0.611 * exp((17.502*temp_mean)/(temp_mean + 240.97)), # saturation VP, Campbell & Normal 1998
         # VPD, Campbell & Normal 1998
         VPD_kPa = e_s_kPa*(1 - (humidity_mean/100))) 
tmt_only_means
```



## Trial Differences

Were there also differences across the trials (experiment iterations)?

```{r}
tmt_trial_means <- format_HOBO_data %>%
  group_by(tmt, trial) %>%
  summarise(temp_mean_tmttrial = mean(temp_C),
            temp_SD_tmttrial = sd(temp_C),
            humidity_mean_tmttrial = mean(relative_humidity),
            humidity_SD_tmttrial = sd(relative_humidity)) %>%
  arrange(trial) %>%
  mutate(e_s_kPa = 0.611 * exp((17.502*temp_mean_tmttrial)/(temp_mean_tmttrial + 240.97)),
         VPD_kPa_tmttrial = e_s_kPa*(1 - (humidity_mean_tmttrial/100)) # Campbell&Norman1998
         )
tmt_trial_means
```


reformat and the mean values:

```{r}
tmt_trial_means$tmt <- factor(tmt_trial_means$tmt,
                           levels = c("cool-humid",
                                      "hot-humid", 
                                      "cool-dry", 
                                      "hot-dry"),
                           labels = c("Cool Humid (0.6 kPa)", 
                                      "Hot Humid (1.1 kPa)", 
                                      "Cool Dry (2.5 kPa)", 
                                      "Hot Dry (3.8 kPa)"))

summary(tmt_trial_means)
write_rds(tmt_trial_means, "./data/HOBO_tmt_trial_diffs.RDS")
write.csv(tmt_trial_means, "./results_statistics/HOBO_mean_diffs.csv")
```




## Models

Run linear models with pairwise post-hoc tests to determine when and how much climate varied among treatments.

```{r}
# temperature
temp_mod <- lm(data = format_HOBO_data, 
               temp_C ~ tmt * trial)
a_temp_mod <- aov(temp_mod)
ph_temp_mod <- TukeyHSD(a_temp_mod)
summary(temp_mod)
temp_mod_df <- data.frame(broom::tidy(ph_temp_mod)) %>%
  arrange(desc(adj.p.value))
write.csv(temp_mod_df, "./results_statistics/HOBO_temp_pairwise_diffs.csv")

# humidity
humidity_mod <- lm(data = format_HOBO_data, 
                   relative_humidity ~ tmt * trial)
a_humid_mod <- aov(humidity_mod)
ph_humid_mod <- TukeyHSD(a_humid_mod)
summary(humidity_mod)
humid_mod_df <- data.frame(broom::tidy(ph_humid_mod)) %>%
  arrange(desc(adj.p.value))
write.csv(humid_mod_df,
          "./results_statistics/HOBO_humidity_pairwise_diffs.csv")
```




## Boxplots

Simple plots looking at the distribution of temperature and humidity for each treatment group:

```{r}
format_HOBO_data %>%
  ggplot() +
  geom_boxplot(aes(x = tmt,
                   y = temp_C,
                   fill = tmt)) +
  theme_classic()
format_HOBO_data %>%
  ggplot() +
  geom_boxplot(aes(x = tmt,
                   y = relative_humidity,
                   fill = tmt)) +
  theme_classic()
```


FANCY boxplot for **temperature**, with ANOVA groups overlaid:

```{r}
# create ANOVA labels
anova_labels_temp <- data.frame(tmt = c("cool-dry",
                                   "cool-humid",
                                   "hot-dry",
                                   "hot-humid"
                                  ),
                                # c-d, c-h, h-d, h-h
                           t1 = c("C", "A", "B", "B"),
                           t2 = c("A", "D", "B", "B"),
                           t3 = c("A", "A", "B", "B"),
                           t4 = c("A", "E", "F", "B"),
                           t5 = c("A", "A", "B", "B"))
# plot!
format_HOBO_data %>%
  ggplot() +
  geom_boxplot(aes(x = trial,
                   y = temp_C,
                   fill = trial)) +
  theme_classic() + 
  facet_wrap(~tmt) +
  xlab("Trial") +
  ylab("Temperature (°C)") +
  ylim(18,38) +
  scale_fill_brewer(palette = "Set2",
                    name = "Trial") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "right") +
  # trial 1 labels
  geom_text(data = anova_labels_temp,
            size = 7,
            mapping = aes(x = 1, y = c(35,35,20,20), label = t1)) +
  # trial 2 labels
  geom_text(data = anova_labels_temp,
            size = 7,
            mapping = aes(x = 2, y = c(35,35,20,20), label = t2)) +
  # trial 3 labels
  geom_text(data = anova_labels_temp,
            size = 7,
            mapping = aes(x = 3, y = c(35,35,20,20), label = t3)) +
  # trial 4 labels
  geom_text(data = anova_labels_temp,
            size = 7,
            mapping = aes(x = 4, y = c(35,35,20,20), label = t4)) +
  # trial 5 labels
  geom_text(data = anova_labels_temp,
            size = 7,
            mapping = aes(x = 5, y = c(35,35,20,20), label = t5)
            ) -> ANOVA_temp_boxplot
ANOVA_temp_boxplot
```


FANCY boxplot for **humidity**, with ANOVA groups overlaid:

```{r}
# create ANOVA labels
anova_labels_humidity <- data.frame(tmt = c("cool-dry",
                                   "cool-humid",
                                   "hot-dry",
                                   "hot-humid"
                                  ),
                                # c-d, c-h, h-d, h-h
                                t1 = c("A", "B", "B", "C"),
                                t2 = c("A", "F", "D", "E"),
                                t3 = c("A", "C", "D", "G"),
                                t4 = c("A", "H", "D", "E"),
                                t5 = c("I", "E", "J", "E"))
# plot!
format_HOBO_data %>%
  ggplot() +
  geom_boxplot(aes(x = trial,
                   y = relative_humidity,
                   fill = trial)) +
  theme_classic() + 
  facet_wrap(~tmt) +
  xlab("Trial") +
  ylab("Relative Humidity (%)") +
  ylim(0,100) +
  scale_fill_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "right") +
  # trial 1 labels
  geom_text(data = anova_labels_humidity,
            size = 7,
            mapping = aes(x = 1, y = c(90,10,90,10), label = t1)) +
  # trial 2 labels
  geom_text(data = anova_labels_humidity,
            size = 7,
            mapping = aes(x = 2, y = c(90,10,90,10), label = t2)) +
  # trial 3 labels
  geom_text(data = anova_labels_humidity,
            size = 7,
            mapping = aes(x = 3, y = c(90,10,90,10), label = t3)) +
  # trial 4 labels
  geom_text(data = anova_labels_humidity,
            size = 7,
            mapping = aes(x = 4, y = c(90,10,90,10), label = t4)) +
  # trial 5 labels
  geom_text(data = anova_labels_humidity,
            size = 7,
            mapping = aes(x = 5, y = c(90,10,90,10), label = t5)
            ) -> ANOVA_humidity_boxplot
ANOVA_humidity_boxplot
```



