---
title: "Climate Water Loss Experiment - Treatment Hydration Analysis"
author: "Savannah Weaver"
date: "2021"
output: pdf_document
toc: TRUE
---

# Packages


```{r setup, include = TRUE, message = FALSE}
`%nin%` = Negate(`%in%`)
if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse") # workflow and plots
if (!require("lme4")) install.packages("lme4")
library("lme4") # for LMMs
if (!require("lmerTest")) install.packages("lmerTest")
library("lmerTest") # for p-values
if (!require("ggpubr")) install.packages("ggpubr")
library("ggpubr") # for multi-ggplot figs
if (!require("UsingR")) install.packages("UsingR")
library("UsingR") # simple.eda model assumption checker
if (!require("broom")) install.packages("broom")
library("broom") # lmer model export
if (!require("broom.mixed")) install.packages("broom.mixed")
library("broom.mixed") # lmer model export
if (!require("car")) install.packages("car")
library("car") # VIFs
if (!require("AICcmodavg")) install.packages("AICcmodavg")
library("AICcmodavg") # model selection
```


# Background and Goals

This data was collected June - August by Master's student Savannah Weaver, advisor Dr. Emily Taylor, and research assistants Tess McIntyre and Taylor Van Rossum. Adult male *Sceloporus occidentalis* were caught across the Cal Poly campus and in Poly Canyon. This R file analyzes the effect of experimental climate treatments on their body condition, osmotic balance, and osmoregulation. Please refer to **doi:** for the published scientific journal article and full details.


# Data

## Load

Read-in data that was compiled, formatted, and checked for completeness in 'capture_analysis'. See that file for information related to the variables. Some variables need to be re-formatted because they lose their data class in export/import. Attach treatment assignments and format that data. Also compute scaled mass index using the equation derived in 'capture_analysis'.

```{r load data}
dat <- read.csv("./data/full_exp_data.csv", # main dataset
                na.strings=c("","NA")) %>%
         # reformat individual_ID to make joining easy
  mutate(individual_ID = as.numeric(individual_ID)) %>%
            # join tmt assignment info
  left_join(read.csv("./data/tmt_assignments.csv"),
            by = "individual_ID") %>%
                # remove irrelevant variables
  dplyr::select(-X, -capture_date, 
                -time_captured, -time_processed, -time_c_temp) %>%
          # format date
  mutate(measurement_date = as.Date(measurement_date,
                                    format = "%Y-%m-%d"),
         # set factor classes
         type = as.factor(type),
         day = as.factor(day),
         individual_ID = as.factor(individual_ID), 
         hemolyzed = as.factor(hemolyzed),
         trial_number = as.factor(trial_number),
         tmt = as.factor(paste(temp_tmt, humidity_tmt)),
         temp_tmt = as.factor(temp_tmt),
         humidity_tmt = as.factor(humidity_tmt),
         conclusion = as.factor(conclusion),
         # set numeric classes
         day_n = as.numeric(day_n),
         mass_g = as.numeric(mass_g),
         hematocrit_percent = as.numeric(hematocrit_percent),
         osmolality_mmol_kg_mean = as.numeric(osmolality_mmol_kg_mean),
         CEWL_g_m2h_mean = as.numeric(CEWL_g_m2h_mean),
         cloacal_temp_C = as.numeric(cloacal_temp_C),
         SVL_mm = as.numeric(SVL_mm),
         # compute scaled mass index
         SMI = mass_g * ((67.71/SVL_mm) ^ 2.65)) %>%
  # only use data from lizards who underwent the full experiment
  # this removes 21 obs for the 7 lizards who didn't complete the exp
  dplyr::filter(conclusion == "complete")
summary(dat)
```



## Clean

There are a handful of points that appear erroneous. 

Individual 239 had post-treatment CEWL >60, which is incredibly unusual for our experiment. He was in the process of shedding when we took this CEWL measurement, so we assume that his process of shedding confounded potential treatment effects, and we should remove that point from our data. We still want to use the other measurements for this individual, so we will just set that CEWL measurement as an NA.

```{r fix erroneous CEWL}
which(dat$CEWL_g_m2h_mean > 60)
dat[332, ]
dat[332, "CEWL_g_m2h_mean"]
dat[332, "CEWL_g_m2h_mean"] <- NA
dat[332, "CEWL_g_m2h_mean"]
dat %>%
  dplyr::filter(complete.cases(CEWL_g_m2h_mean)) %>%
  summarise(max(CEWL_g_m2h_mean))
```


10 trial 1 lizards have unreasonably high osmolality measurements on June 24, which we think are due to an osmometer technical error. The values are way too far outside the usual range to be trustworthy, so we will exclude them.

```{r fix erroneous osml}
which(dat$osmolality_mmol_kg_mean > 500)
dat[136:145, ]
dat[136:145, "osmolality_mmol_kg_mean"]
dat[136:145, "osmolality_mmol_kg_mean"] <- c(rep(NA, 10))
dat[136:145, "osmolality_mmol_kg_mean"]
dat[136:145, ]
dat %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  summarise(max(osmolality_mmol_kg_mean))
```




## Format

Rename some factors:

```{r rename factors}
dat$humidity_tmt <- factor(dat$humidity_tmt,
                           levels = c("humid", "dry"),
                           labels = c("Humid", "Dry"))
dat$temp_tmt <- factor(dat$temp_tmt,
                           levels = c("hot", "cool"),
                           labels = c("Hot", "Cool"))
dat$tmt <- factor(dat$tmt,
                           levels = c("hot humid", "hot dry",
                                      "cool humid", "cool dry"),
                           labels = c("Hot Humid", "Hot Dry",
                                      "Cool Humid", "Cool Dry"))
summary(dat)
```


Make sub-dataframes without rehab data / with only rehab-related data:

```{r sub-dfs}
dat_no_rehab <- dat %>%
  dplyr::filter(type == "exp")
dat_for_rehab <- dat %>%
  dplyr::filter(day_n %in% c(8, 10))
```



## Check

Dates:

```{r check dates}
unique(dat$measurement_date)
```

Number of measurements for each lizard:

```{r check n msmts}
dat_no_rehab %>%
  group_by(individual_ID) %>%
  summarise(n = n()) %>%
  arrange(n)
```


Every lizard has 6 experimental measurements: pre-tmt, mid-tmt, post-tmt, and mass checks on each of the 3 days between mid and post-tmt.


Did any of the treatment groups inherently start out with large differences in response variables?


```{r check starting diffs}
dat %>%
  dplyr::filter(day == "capture") %>%
  group_by(tmt) %>%
  summarise(mean(mass_g),
            mean(SMI),
            mean(hematocrit_percent),
            mean(osmolality_mmol_kg_mean),
            mean(CEWL_g_m2h_mean))
```

There are slight differences, but overall the starting values across groups are more or less the same.



# Experiment Models

We predicted that there would be effects of day, humidity, and temperature, but we only expect those interactions with day to have an effect because we're observing changes over time.

So, each full model will include the following: day_n, day_n:humidity_tmt, day_n:temp_tmt, and day_n:humidity_tmt:temp_tmt. Then I'll test each possible sub-combination

## SMI

### Building

Sort based on number of predictors:

#### 4

```{r SMI exp model 4}
SMI_mod4 <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ day_n + 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(SMI_mod4) # check collinearity 
```

#### 3a

```{r SMI exp model 3a}
SMI_mod3a <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                         day_n + 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         (1|trial_number/individual_ID))
car::vif(SMI_mod3a) # check collinearity 
```

#### 3b

```{r SMI exp model 3b}
SMI_mod3b <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ day_n + 
                         day_n:humidity_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(SMI_mod3b) # check collinearity 
```

#### 3c

```{r SMI exp model 3c}
SMI_mod3c <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ day_n + 
                         day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(SMI_mod3c) # check collinearity 
```

#### 3d

```{r SMI exp model 3d}
SMI_mod3d <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(SMI_mod3d) # check collinearity 
```


#### 2a

```{r SMI exp model 2a}
SMI_mod2a <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ day_n + 
                         day_n:humidity_tmt + 
                         (1|trial_number/individual_ID))
```

#### 2b

```{r SMI exp model 2b}
SMI_mod2b <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ day_n + 
                         day_n:temp_tmt +   
                         (1|trial_number/individual_ID))
```

#### 2c

```{r SMI exp model 2c}
SMI_mod2c <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ day_n +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
```

#### 2d

```{r SMI exp model 2d}
SMI_mod2d <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         (1|trial_number/individual_ID))
```

#### 2e

```{r SMI exp model 2e}
SMI_mod2e <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                         day_n:humidity_tmt + 
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
```

#### 2f

```{r SMI exp model 2f}
SMI_mod2f <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                         day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
```

#### 1

single-predictor effects
                   
```{r SMI exp model 1 variables}
SMI_mod1a <- lme4::lmer(data = dat_no_rehab, SMI ~ 
                          day_n +
                         (1|trial_number/individual_ID))
SMI_mod1b <- lme4::lmer(data = dat_no_rehab, SMI ~ day_n + 
                         day_n:humidity_tmt + 
                         (1|trial_number/individual_ID))
SMI_mod1c <- lme4::lmer(data = dat_no_rehab, SMI ~ 
                          day_n:temp_tmt + 
                         (1|trial_number/individual_ID))
SMI_mod1d <- lme4::lmer(data = dat_no_rehab, SMI ~
                          day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
```

#### null model

```{r SMI exp model NULL}
SMI_mod_null <- lme4::lmer(data = dat_no_rehab, SMI ~ 1 +
                         (1|trial_number/individual_ID))
```


### Selection

Compare all models:

```{r SMI mod compare}
SMI_models <- list(SMI_mod4, 
                   SMI_mod3a, SMI_mod3b, SMI_mod3c, SMI_mod3d, 
                   SMI_mod2a, SMI_mod2b, SMI_mod2c,
                   SMI_mod2d, SMI_mod2e, SMI_mod2f,
                   SMI_mod1a, SMI_mod1b, SMI_mod1c, SMI_mod1d,
                   SMI_mod_null)

#specify model names
EXP_mod_names <- data.frame(predictors = 
                 c('Day, Day:Humd, Day:Temp, Day:Humd:Temp', #4
                   'Day, Day:Humd, Day:Temp', #3a
                   'Day, Day:Humd, Day:Humd:Temp', #3b
                   'Day, Day:Temp, Day:Humd:Temp', #3c
                   'Day:Humd, Day:Temp, Day:Humd:Temp', #3d
                   'Day, Day:Humd', #2a
                   'Day, Day:Temp', #2b
                   'Day, Day:Humd:Temp', #2c
                   'Day:Humd, Day:Temp', #2d
                   'Day:Humd, Day:Humd:Temp', #2e
                   'Day:Temp, Day:Humd:Temp', #2f
                   'Day', #1a
                   'Day:Humd', #1b
                   'Day:Temp', #1c
                   'Day:Humd:Temp', #1d
                   'Null'))
rownames(EXP_mod_names) <- c('4', '3a', '3b', '3c', '3d',
                             '2a', '2b', '2c', '2d', '2e', '2f',
                             '1a', '1b', '1c', '1d', 'null')

#calculate AIC of each model
SMI_AICc <- data.frame(aictab(cand.set = SMI_models,
                              modnames = EXP_mod_names)) %>%
  mutate(response = "Body Condition (g)")
SMI_AICc
```


SMI is explained equally well by the top two models: 2d and 3a

### Assumptions

Check linear regression assumptions/conditions:

```{r SMI exp model assumptions}
plot(SMI_mod2d)
simple.eda(residuals(SMI_mod2d))
shapiro.test(residuals(SMI_mod2d))
plot(SMI_mod3a)
simple.eda(residuals(SMI_mod3a))
shapiro.test(residuals(SMI_mod3a))
```

Normality is very violated, but equal error variance and linearity look fantastic, so I think it's okay.


### Save

Save best model stats with p-values:

```{r best SMI exp model}
SMI_mod_best1 <- lmerTest::lmer(data = dat_no_rehab, 
                                SMI ~ 
                                 day_n:humidity_tmt + day_n:temp_tmt +
                                 (1|trial_number/individual_ID))
summary(SMI_mod_best1)

SMI_mod_best2 <- lmerTest::lmer(data = dat_no_rehab, 
                                SMI ~ 
                                 day_n + 
                                  day_n:humidity_tmt + day_n:temp_tmt +
                                 (1|trial_number/individual_ID))
summary(SMI_mod_best2)

SMI_stats <- broom.mixed::tidy(SMI_mod_best1) %>%
  rbind(broom.mixed::tidy(SMI_mod_best2)) %>%
  mutate(response = "Body Condition (g)")
```



## Hematocrit

### Building

#### 4

```{r hct exp model 4}
hct_mod4 <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ day_n + 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(hct_mod4) # check collinearity 
```


#### 3a

```{r hct exp model 3a}
hct_mod3a <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         day_n + 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         (1|trial_number/individual_ID))
car::vif(hct_mod3a) # check collinearity 
```

#### 3b

```{r hct exp model 3b}
hct_mod3b <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         day_n + 
                         day_n:humidity_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(hct_mod3b) # check collinearity 
```

#### 3c

```{r hct exp model 3c}
hct_mod3c <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ day_n + 
                         day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(hct_mod3c) # check collinearity 
```

#### 3d

```{r hct exp model 3d}
hct_mod3d <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(hct_mod3d) # check collinearity 
```

#### 2a

```{r hct exp model 2a}
hct_mod2a <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ day_n + 
                         day_n:humidity_tmt + 
                         (1|trial_number/individual_ID))
car::vif(hct_mod2a) # check collinearity 
```

#### 2b

```{r hct exp model 2b}
hct_mod2b <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ day_n + 
                         day_n:temp_tmt +   
                         (1|trial_number/individual_ID))
car::vif(hct_mod2b) # check collinearity 
```

#### 2c

```{r hct exp model 2c}
hct_mod2c <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ day_n +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(hct_mod2c) # check collinearity 
```

#### 2d

```{r hct exp model 2d}
hct_mod2d <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         (1|trial_number/individual_ID))
car::vif(hct_mod2d) # check collinearity 
```


#### 2e

```{r hct exp model 2e}
hct_mod2e <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         day_n:humidity_tmt + 
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(hct_mod2e) # check collinearity 
```

#### 2f

```{r hct exp model 2f}
hct_mod2f <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(hct_mod2f) # check collinearity 
```


#### 1
test each single effect model:
                   
```{r hct exp model 1s}
hct_mod1a <- lme4::lmer(data = dat_no_rehab, 
                        hematocrit_percent ~ 
                          day_n +
                         (1|trial_number/individual_ID))
hct_mod1b <- lme4::lmer(data = dat_no_rehab, 
                        hematocrit_percent ~ 
                          day_n:humidity_tmt +
                         (1|trial_number/individual_ID))
hct_mod1c <- lme4::lmer(data = dat_no_rehab, 
                        hematocrit_percent ~ 
                          day_n:temp_tmt +
                         (1|trial_number/individual_ID))
hct_mod1d <- lme4::lmer(data = dat_no_rehab, 
                        hematocrit_percent ~
                          day_n:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
```

#### null model

```{r hct exp model NULL}
hct_mod_null <- lme4::lmer(data = dat_no_rehab, hematocrit_percent ~ 1 +
                         (1|trial_number/individual_ID))
```


### Selection

Compare all models:

```{r hct mod compare}
hct_models <- list(hct_mod4, 
                   hct_mod3a, hct_mod3b, hct_mod3c, hct_mod3d, 
                   hct_mod2a, hct_mod2b, hct_mod2c,
                   hct_mod2d, hct_mod2e, hct_mod2f,
                   hct_mod1a, hct_mod1b, hct_mod1c, hct_mod1d,
                   hct_mod_null)

#calculate AIC of each model
hct_AICc <- data.frame(aictab(cand.set = hct_models, 
                                 modnames = EXP_mod_names)) %>%
  mutate(response = "Hematocrit (%)")
hct_AICc
```


Explained equally well by the top 2 models: 1c (Day:Temp) and 2b (Day, Day:Temp)


### Assumptions

Check linear regression assumptions/conditions:

```{r hct exp model assumptions}
plot(hct_mod1c)
simple.eda(residuals(hct_mod1c))
shapiro.test(residuals(hct_mod1c))
plot(hct_mod2b)
simple.eda(residuals(hct_mod2b))
shapiro.test(residuals(hct_mod2b))
```

All assumptions are satisfied (and beautiful). 



### Save

Save best model stats (1c, 2b) with p-values:

```{r best hct exp models}
hct_mod_best1 <- lmerTest::lmer(data = dat_no_rehab,
                        hematocrit_percent ~ 
                         day_n:temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_mod_best1)

hct_mod_best2 <- lmerTest::lmer(data = dat_no_rehab,
                        hematocrit_percent ~ 
                         day_n + day_n:temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_mod_best2)

hct_stats <- broom.mixed::tidy(hct_mod_best1) %>%
  rbind(broom.mixed::tidy(hct_mod_best2)) %>%
  mutate(response = "Hematocrit (%)")
```







## Osmolality


### Building

#### 4

```{r osml exp model 4}
osml_mod4 <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ day_n + 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(osml_mod4) # check collinearity 
```


#### 3a

```{r osml exp model 3a}
osml_mod3a <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         day_n + 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         (1|trial_number/individual_ID))
car::vif(osml_mod3a) # check collinearity 
```

#### 3b

```{r osml exp model 3b}
osml_mod3b <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ day_n + 
                         day_n:humidity_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(osml_mod3b) # check collinearity 
```

#### 3c

```{r osml exp model 3c}
osml_mod3c <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ day_n + 
                         day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(osml_mod3c) # check collinearity 
```

#### 3d

```{r osml exp model 3d}
osml_mod3d <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(osml_mod3d) # check collinearity 
```

#### 2a

```{r osml exp model 2a}
osml_mod2a <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ day_n + 
                         day_n:humidity_tmt + 
                         (1|trial_number/individual_ID))
car::vif(osml_mod2a) # check collinearity 
```

#### 2b

```{r osml exp model 2b}
osml_mod2b <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ day_n + 
                         day_n:temp_tmt +   
                         (1|trial_number/individual_ID))
car::vif(osml_mod2b) # check collinearity 
```

#### 2c

```{r osml exp model 2c}
osml_mod2c <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ day_n +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(osml_mod2c) # check collinearity 
```


#### 2d

```{r osml exp model 2d}
osml_mod2d <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         (1|trial_number/individual_ID))
car::vif(osml_mod2d) # check collinearity 
```

#### 2e

```{r osml exp model 2e}
osml_mod2e <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         day_n:humidity_tmt + 
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(osml_mod2e) # check collinearity 
```

#### 2f

```{r osml exp model 2f}
osml_mod2f <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(osml_mod2f) # check collinearity 
```




#### 1

```{r osml exp model 1}
osml_mod1a <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         day_n +
                         (1|trial_number/individual_ID))
osml_mod1b <- lme4::lmer(data = dat_no_rehab, 
                         osmolality_mmol_kg_mean ~ 
                           day_n:humidity_tmt +
                         (1|trial_number/individual_ID))
osml_mod1c <- lme4::lmer(data = dat_no_rehab, 
                         osmolality_mmol_kg_mean ~ 
                           day_n:temp_tmt +
                         (1|trial_number/individual_ID))
osml_mod1d <- lme4::lmer(data = dat_no_rehab, 
                         osmolality_mmol_kg_mean ~ 
                           day_n:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
```


#### null

```{r osml exp model NULL}
osml_mod_null <- lme4::lmer(data = dat_no_rehab, 
                            osmolality_mmol_kg_mean ~ 1 +
                         (1|trial_number/individual_ID))
```


### Selection

Compare all models:

```{r osml mod compare}
osml_models <- list(osml_mod4, 
                   osml_mod3a, osml_mod3b, osml_mod3c, osml_mod3d, 
                   osml_mod2a, osml_mod2b, osml_mod2c,
                   osml_mod2d, osml_mod2e, osml_mod2f,
                   osml_mod1a, osml_mod1b, osml_mod1c, osml_mod1d,
                   osml_mod_null)

#calculate AIC of each model
osml_AICc <- data.frame(aictab(cand.set = osml_models, 
                                 modnames = EXP_mod_names)) %>%
  mutate(response = "Plasma Osmolality (mmol/kg)")
osml_AICc
```


... ?all? the models are best model?! Ity seems like 2e is the fewest variables with the most significant relationships, so we will present that one.


### Assumptions

Check linear regression assumptions/conditions:

```{r osml exp model assumptions}
plot(osml_mod2e)
simple.eda(residuals(osml_mod2e))
shapiro.test(residuals(osml_mod2e))
```

All assumptions are OKAY.



### Save

Save best model stats with p-values:

```{r best osml exp models}
osml_mod_best1 <- lmerTest::lmer(data = dat_no_rehab,
                        osmolality_mmol_kg_mean ~ 
                         day_n:humidity_tmt + 
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
summary(osml_mod_best1)
osml_stats <- broom.mixed::tidy(osml_mod_best1) %>%
  mutate(response = "Plasma Osmolality (mmol/kg)")
```




## CEWL


### Building

#### 4

```{r CEWL exp model 4}
dat_no_rehab_rm <- dat_no_rehab %>%
  dplyr::filter(complete.cases(.))
CEWL_mod4 <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C + 
                         day_n + 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod4) # check collinearity 
```


#### 3a

```{r CEWL exp model 3a}
CEWL_mod3a <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C +  
                         day_n + 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod3a) # check collinearity 
```

#### 3b

```{r CEWL exp model 3b}
CEWL_mod3b <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C +  day_n + 
                         day_n:humidity_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod3b) # check collinearity 
```

#### 3c

```{r CEWL exp model 3c}
CEWL_mod3c <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C +  day_n + 
                         day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod3c) # check collinearity 
```

#### 3d

```{r CEWL exp model 3d}
CEWL_mod3d <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C +  
                         day_n:humidity_tmt + day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod3d) # check collinearity 
```

#### 2a

```{r CEWL exp model 2a}
CEWL_mod2a <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C +  day_n + 
                         day_n:humidity_tmt + 
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod2a) # check collinearity 
```

#### 2b

```{r CEWL exp model 2b}
CEWL_mod2b <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C +  day_n + 
                         day_n:temp_tmt +   
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod2b) # check collinearity 
```

#### 2c

```{r CEWL exp model 2c}
CEWL_mod2c <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C +  day_n +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod2c) # check collinearity 
```


#### 2d

```{r CEWL exp model 2d}
CEWL_mod2d <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C +  
                         day_n:humidity_tmt + day_n:temp_tmt +
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod2d) # check collinearity 
```

#### 2e

```{r CEWL exp model 2e}
CEWL_mod2e <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C +  
                         day_n:humidity_tmt + 
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod2e) # check collinearity 
```

#### 2f

```{r CEWL exp model 2f}
CEWL_mod2f <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C +  
                         day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod2f) # check collinearity 
```




#### 1

```{r CEWL exp model 1}
CEWL_mod1a <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C +  
                         day_n +
                         (1|trial_number/individual_ID))
CEWL_mod1b <- lme4::lmer(data = dat_no_rehab_rm, 
                         CEWL_g_m2h_mean ~ cloacal_temp_C +  
                           day_n:humidity_tmt +
                         (1|trial_number/individual_ID))
CEWL_mod1c <- lme4::lmer(data = dat_no_rehab_rm, 
                         CEWL_g_m2h_mean ~ cloacal_temp_C +  
                           day_n:temp_tmt +
                         (1|trial_number/individual_ID))
CEWL_mod1d <- lme4::lmer(data = dat_no_rehab_rm, 
                         CEWL_g_m2h_mean ~ cloacal_temp_C +  
                           day_n:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
```


#### null

```{r CEWL exp model NULL}
CEWL_mod_null <- lme4::lmer(data = dat_no_rehab_rm, 
                            CEWL_g_m2h_mean ~ cloacal_temp_C +  1 +
                         (1|trial_number/individual_ID))
```





### Selection

Compare all models:

```{r CEWL mod compare}
CEWL_models <- list(CEWL_mod4, 
                   CEWL_mod3a, CEWL_mod3b, CEWL_mod3c, CEWL_mod3d, 
                   CEWL_mod2a, CEWL_mod2b, CEWL_mod2c,
                   CEWL_mod2d, CEWL_mod2e, CEWL_mod2f,
                   CEWL_mod1a, CEWL_mod1b, CEWL_mod1c, CEWL_mod1d,
                   CEWL_mod_null)
CEWL_mod_names <- EXP_mod_names %>%
  mutate(predictors = paste(predictors, "BodyTemp", sep = ", "))

#calculate AIC of each model
CEWL_AICc <- data.frame(aictab(cand.set = CEWL_models, 
                                 modnames = CEWL_mod_names)) %>%
  mutate(response = "CEWL (g/m2h)")
CEWL_AICc
```

Predicted equally well by the top 8 models...

Every variable in the full model has a significant relationship, so we will present that model.


### Assumptions

Check linear regression assumptions/conditions:

```{r CEWL exp model assumptions}
plot(CEWL_mod4)
simple.eda(residuals(CEWL_mod4))
shapiro.test(residuals(CEWL_mod4))
```

normality is not good, both other assumptions OKAY enough.



### Save

Save best model stats with p-values:

```{r best CEWL exp model}
CEWL_mod_best <- lmerTest::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ cloacal_temp_C + 
                         day_n + 
                         day_n:humidity_tmt + day_n:temp_tmt +
                         day_n:humidity_tmt:temp_tmt + 
                         (1|trial_number/individual_ID))
summary(CEWL_mod_best)
CEWL_stats <- broom.mixed::tidy(CEWL_mod_best) %>%
  mutate(response = "CEWL (g/m2h)")
```



## Group Export

Put all the best models and AICc rankings into one csv each

```{r export exp stats}
experiment_stats <- CEWL_stats %>%
  rbind(osml_stats) %>%
  rbind(hct_stats) %>%
  rbind(SMI_stats)
write.csv(experiment_stats, 
          "./results_statistics/exp_best_models.csv")

AICc_ranks <- CEWL_AICc %>%
  rbind(osml_AICc) %>%
  rbind(hct_AICc) %>%
  rbind(SMI_AICc)
write.csv(AICc_ranks, "./results_statistics/exp_model_rankings.csv")
```







# Rehydration Models


## SMI

### Building

```{r SMI rehydration model 1}
SMI_rmod1 <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ 
                         type*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(SMI_rmod1)
drop1(SMI_rmod1)
```

Remove three-way interaction and day*humidity interaction.

```{r SMI rehydration model 2}
SMI_rmod2 <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ 
                         type*humidity_tmt*temp_tmt -
                         type:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
summary(SMI_rmod2)
drop1(SMI_rmod2)
```


try each 2-interaction model:

```{r SMI rehydration model 2abc}
SMI_rmod2a <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ type*(humidity_tmt + temp_tmt) +
                         (1|trial_number/individual_ID))
SMI_rmod2b <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ (type + humidity_tmt)*temp_tmt +
                         (1|trial_number/individual_ID))
SMI_rmod2c <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ (type + temp_tmt) * humidity_tmt +
                         (1|trial_number/individual_ID))
```

Drop to only one interaction:

```{r SMI rehydration model 3abc}
SMI_rmod3a <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ type*humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
SMI_rmod3b <- lme4::lmer(data = dat_for_rehab,
                       SMI ~  humidity_tmt + type*temp_tmt +
                         (1|trial_number/individual_ID))
SMI_rmod3c <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ type + temp_tmt * humidity_tmt +
                         (1|trial_number/individual_ID))
```

No interactions:

```{r SMI rehydration model 4}
SMI_rmod4 <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ type + temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
```


try each 2-variable combination:

```{r SMI rehydration model 4abc}
SMI_rmod4a <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ type + humidity_tmt +
                         (1|trial_number/individual_ID))
SMI_rmod4b <- lme4::lmer(data = dat_for_rehab,
                       SMI ~  type + temp_tmt +
                         (1|trial_number/individual_ID))
SMI_rmod4c <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
```

try each single variable:

```{r SMI rehydration model 5abc}
SMI_rmod5a <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ type +
                         (1|trial_number/individual_ID))
SMI_rmod5b <- lme4::lmer(data = dat_for_rehab,
                       SMI ~  temp_tmt +
                         (1|trial_number/individual_ID))
SMI_rmod5c <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ humidity_tmt +
                         (1|trial_number/individual_ID))
```


Lastly, null model:


```{r SMI rehydration model null}
SMI_rmod_null <- lme4::lmer(data = dat_for_rehab,
                            SMI ~ 1 + (1|trial_number/individual_ID))
```




### Selection

Compare all models:

```{r SMI rehydration mod compare}
SMI_rmodels <- list(SMI_rmod1, 
                   SMI_rmod2, SMI_rmod2a, SMI_rmod2b, SMI_rmod2c, 
                              SMI_rmod3a, SMI_rmod3b, SMI_rmod3c, 
                   SMI_rmod4, SMI_rmod4a, SMI_rmod4b, SMI_rmod4c, 
                              SMI_rmod5a, SMI_rmod5b, SMI_rmod5c, 
                   SMI_rmod_null)

#specify model names
REHYD_mod_names <- c('(model 1) ~ Rhyd, Humd, Temp, Rhyd:Humd, Rhyd:Temp, Humd:Temp, Rhyd:Humd:Temp',
                   '(model 2) ~ Rhyd, Humd, Temp, Rhyd:Humd, Rhyd:Temp, Humd:Temp',
                   '(model 2a) ~ Rhyd, Humd, Temp, Rhyd:Humd, Rhyd:Temp',
                   '(model 2b) ~ Rhyd, Humd, Temp, Rhyd:Temp, Humd:Temp',
                   '(model 2c) ~ Rhyd, Humd, Temp, Rhyd:Humd, Humd:Temp',
                   '(model 3a) ~ Rhyd, Humd, Temp, Rhyd:Humd',
                   '(model 3b) ~ Rhyd, Humd, Temp, Rhyd:Temp',
                   '(model 3c) ~ Rhyd, Humd, Temp, Humd:Temp',
                   '(model 4) ~ Rhyd, Humd, Temp',
                   '(model 4a) ~ Rhyd, Humd',
                   '(model 4b) ~ Rhyd, Temp',
                   '(model 4c) ~ Humd, Temp',
                   '(model 5a) ~ Rhyd',
                   '(model 5b) ~ Temp',
                   '(model 5c) ~ Humd',
                   'null model')

#calculate AIC of each model
SMI_rAICc <- data.frame(aictab(cand.set = SMI_rmodels, 
                                 modnames = REHYD_mod_names))
SMI_rAICc
```


Models 3b and 4b tied for best model.


### Assumptions

Check linear regression assumptions/conditions:

```{r SMI rehydration model assumptions}
plot(SMI_rmod4b)
simple.eda(residuals(SMI_rmod4b))
shapiro.test(residuals(SMI_rmod4b))
plot(SMI_rmod3b)
simple.eda(residuals(SMI_rmod3b))
shapiro.test(residuals(SMI_rmod3b))
```

All of the assumptions are met very well.


### Export

Save best model stats with p-values:

```{r best SMI rehydration models}
SMI_rmod_best1 <- lmerTest::lmer(data = dat_for_rehab,
                       SMI ~  type + temp_tmt +
                         (1|trial_number/individual_ID))
summary(SMI_rmod_best1)
write.csv(broom.mixed::tidy(SMI_rmod_best1), 
          "./results_statistics/rehydration_SMI_best_mod1.csv")
SMI_rmod_best2 <- lmerTest::lmer(data = dat_for_rehab,
                       SMI ~  humidity_tmt + type*temp_tmt +
                         (1|trial_number/individual_ID))
summary(SMI_rmod_best2)
write.csv(broom.mixed::tidy(SMI_rmod_best2), 
          "./results_statistics/rehydration_SMI_best_mod2.csv")
write.csv(SMI_rAICc, "./results_statistics/rehydration_SMI_mod_rankings.csv")
```







## Hematocrit



### Building

```{r hct rehydration model 1}
hct_rmod1 <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ 
                         type*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_rmod1)
drop1(hct_rmod1)
```

Remove three-way interaction and day*humidity interaction.

```{r hct rehydration model 2}
hct_rmod2 <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ 
                         type*humidity_tmt*temp_tmt -
                         type:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_rmod2)
drop1(hct_rmod2)
```


try each 2-interaction model:

```{r hct rehydration model 2abc}
hct_rmod2a <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ type*(humidity_tmt + temp_tmt) +
                         (1|trial_number/individual_ID))
hct_rmod2b <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ (type + humidity_tmt)*temp_tmt +
                         (1|trial_number/individual_ID))
hct_rmod2c <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ (type + temp_tmt) * humidity_tmt +
                         (1|trial_number/individual_ID))
```

Drop to only one interaction:

```{r hct rehydration model 3abc}
hct_rmod3a <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ type*humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
hct_rmod3b <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~  humidity_tmt + type*temp_tmt +
                         (1|trial_number/individual_ID))
hct_rmod3c <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ type + temp_tmt * humidity_tmt +
                         (1|trial_number/individual_ID))
```

No interactions:

```{r hct rehydration model 4}
hct_rmod4 <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ type + temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
```


try each 2-variable combination:

```{r hct rehydration model 4abc}
hct_rmod4a <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ type + humidity_tmt +
                         (1|trial_number/individual_ID))
hct_rmod4b <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~  type + temp_tmt +
                         (1|trial_number/individual_ID))
hct_rmod4c <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
```

try each single variable:

```{r hct rehydration model 5abc}
hct_rmod5a <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ type +
                         (1|trial_number/individual_ID))
hct_rmod5b <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~  temp_tmt +
                         (1|trial_number/individual_ID))
hct_rmod5c <- lme4::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ humidity_tmt +
                         (1|trial_number/individual_ID))
```


Lastly, null model:


```{r hct rehydration model null}
hct_rmod_null <- lme4::lmer(data = dat_for_rehab,
                            hematocrit_percent ~ 1 + (1|trial_number/individual_ID))
```




### Selection

Compare all models:

```{r hct rehydration mod compare}
hct_rmodels <- list(hct_rmod1, 
                   hct_rmod2, hct_rmod2a, hct_rmod2b, hct_rmod2c, 
                              hct_rmod3a, hct_rmod3b, hct_rmod3c, 
                   hct_rmod4, hct_rmod4a, hct_rmod4b, hct_rmod4c, 
                              hct_rmod5a, hct_rmod5b, hct_rmod5c, 
                   hct_rmod_null)

#calculate AIC of each model
hct_rAICc <- data.frame(aictab(cand.set = hct_rmodels, 
                                 modnames = REHYD_mod_names))
hct_rAICc
```


There are 3 top models: 1, 2b, and 2


### Assumptions

Check linear regression assumptions/conditions:

```{r hct rehydration model assumptions}
plot(hct_rmod1)
simple.eda(residuals(hct_rmod1))
shapiro.test(residuals(hct_rmod1))
plot(hct_rmod2b)
simple.eda(residuals(hct_rmod2b))
shapiro.test(residuals(hct_rmod2b))
plot(hct_rmod2)
simple.eda(residuals(hct_rmod2))
shapiro.test(residuals(hct_rmod2))
```

conditions are all acceptable


### Export

Save best model stats with p-values:

```{r best hct rehydration models}
# model 1
hct_rmod_best1 <- lmerTest::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ type*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_rmod_best1)
write.csv(broom.mixed::tidy(hct_rmod_best1), 
          "./results_statistics/rehydration_hct_best_mod1.csv")
# model 2b
hct_rmod_best2 <- lmerTest::lmer(data = dat_for_rehab,
                         hematocrit_percent ~ 
                           (type + humidity_tmt)*temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_rmod_best2)
write.csv(broom.mixed::tidy(hct_rmod_best2), 
          "./results_statistics/rehydration_hct_best_mod2.csv")
# model 2
hct_rmod_best3 <- lmerTest::lmer(data = dat_for_rehab,
                       hematocrit_percent ~ 
                         type*humidity_tmt*temp_tmt -
                         type:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_rmod_best3)
write.csv(broom.mixed::tidy(hct_rmod_best3), 
          "./results_statistics/rehydration_hct_best_mod3.csv")
write.csv(hct_rAICc, "./results_statistics/rehydration_hct_mod_rankings.csv")
```




## Osmolality


### Building

```{r osml rehydration model 1}
osml_rmod1 <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ 
                         type*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_rmod1)
drop1(osml_rmod1)
```

Remove three-way interaction and day*humidity interaction.

```{r osml rehydration model 2}
osml_rmod2 <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ 
                         type*humidity_tmt*temp_tmt -
                         type:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_rmod2)
drop1(osml_rmod2)
```


try each 2-interaction model:

```{r osml rehydration model 2abc}
osml_rmod2a <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ type*(humidity_tmt + temp_tmt) +
                         (1|trial_number/individual_ID))
osml_rmod2b <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ (type + humidity_tmt)*temp_tmt +
                         (1|trial_number/individual_ID))
osml_rmod2c <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ (type + temp_tmt) * humidity_tmt +
                         (1|trial_number/individual_ID))
```

Drop to only one interaction:

```{r osml rehydration model 3abc}
osml_rmod3a <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ type*humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
osml_rmod3b <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~  humidity_tmt + type*temp_tmt +
                         (1|trial_number/individual_ID))
osml_rmod3c <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ type + temp_tmt * humidity_tmt +
                         (1|trial_number/individual_ID))
```

No interactions:

```{r osml rehydration model 4}
osml_rmod4 <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ type + temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
```


try each 2-variable combination:

```{r osml rehydration model 4abc}
osml_rmod4a <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ type + humidity_tmt +
                         (1|trial_number/individual_ID))
osml_rmod4b <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~  type + temp_tmt +
                         (1|trial_number/individual_ID))
osml_rmod4c <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
```

try each single variable:

```{r osml rehydration model 5abc}
osml_rmod5a <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ type +
                         (1|trial_number/individual_ID))
osml_rmod5b <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~  temp_tmt +
                         (1|trial_number/individual_ID))
osml_rmod5c <- lme4::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~ humidity_tmt +
                         (1|trial_number/individual_ID))
```


Lastly, null model:


```{r osml rehydration model null}
osml_rmod_null <- lme4::lmer(data = dat_for_rehab,
                            osmolality_mmol_kg_mean ~ 1 +
                              (1|trial_number/individual_ID))
```




### Selection

Compare all models:

```{r osml rehydration mod compare}
osml_rmodels <- list(osml_rmod1, 
                   osml_rmod2, osml_rmod2a, osml_rmod2b, osml_rmod2c, 
                              osml_rmod3a, osml_rmod3b, osml_rmod3c, 
                   osml_rmod4, osml_rmod4a, osml_rmod4b, osml_rmod4c, 
                              osml_rmod5a, osml_rmod5b, osml_rmod5c, 
                   osml_rmod_null)

#calculate AIC of each model
osml_rAICc <- data.frame(aictab(cand.set = osml_rmodels, 
                                 modnames = REHYD_mod_names))
osml_rAICc
```


The full model is the single best model.


### Assumptions

Check linear regression assumptions/conditions:

```{r osml rehydration model assumptions}
plot(osml_rmod1)
simple.eda(residuals(osml_rmod1))
shapiro.test(residuals(osml_rmod1))
```

These are acceptable.


### Export

Save best model stats with p-values:

```{r best osml rehydration model}
osml_rmod_best1 <- lmerTest::lmer(data = dat_for_rehab,
                       osmolality_mmol_kg_mean ~
                         humidity_tmt*type*temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_rmod_best1)
write.csv(broom.mixed::tidy(osml_rmod_best1), 
          "./results_statistics/rehydration_osml_best_mod.csv")
write.csv(osml_rAICc, "./results_statistics/rehydration_osml_mod_rankings.csv")
```









# Figures

## Calculate 

### Predicted Values

```{r mod predicts for plots}
# exp SMI
exp_predict_SMI <- dat_no_rehab %>%
  dplyr::filter(complete.cases(SMI)) %>%
  mutate(SMI_est = predict(SMI_mod_best))
# exp hct
exp_predict_hct <- dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  mutate(hct_est1 = predict(hct_mod_best1),
         hct_est2 = predict(hct_mod_best2),
         hct_est3 = predict(hct_mod_best3))
# exp osml
exp_predict_osml <- dat_no_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  mutate(osml_est1 = predict(osml_mod_best1),
         osml_est2 = predict(osml_mod_best2),
         osml_est3 = predict(osml_mod_best3))
# exp CEWL
exp_predict_CEWL <- dat_no_rehab %>%
  dplyr::filter(complete.cases(CEWL_g_m2h_mean)) %>%
  mutate(CEWL_est = predict(CEWL_mod_best))
# rehydration SMI
rhyd_predict_SMI <- dat_for_rehab %>%
  dplyr::filter(complete.cases(SMI)) %>%
  mutate(SMI_est1 = predict(SMI_rmod_best1),
         SMI_est2 = predict(SMI_rmod_best2)) #%>%
  #dplyr::filter(day_n == 10)
# rehydration hct
rhyd_predict_hct <- dat_for_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  mutate(hct_est1 = predict(hct_rmod_best1),
         hct_est2 = predict(hct_rmod_best2),
         hct_est3 = predict(hct_rmod_best3)) #%>%
  #dplyr::filter(day_n == 10)
# rehydration osml
rhyd_predict_osml <- dat_for_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  mutate(osml_est = predict(osml_rmod_best))# %>%
  #dplyr::filter(day_n == 10)
```



### Means to Overlay

Predicted value means:

```{r predicted means}
# exp SMI
exp_predict_SMI_means <- exp_predict_SMI %>%
  dplyr::filter(day_n %in% c(0, 8)) %>%
  group_by(day_n, tmt) %>%
  summarise(mean_SMI_est = mean(SMI_est))
# exp hct
exp_predict_hct_means <- exp_predict_hct %>%
  dplyr::filter(day_n %in% c(0, 8)) %>%
  group_by(day_n, tmt) %>%
  summarise(mean_hct_est1 = mean(hct_est1),
            mean_hct_est2 = mean(hct_est2),
            mean_hct_est3 = mean(hct_est3))
# exp osml
exp_predict_osml_means <- exp_predict_osml %>%
  dplyr::filter(day_n %in% c(0, 8)) %>%
  group_by(day_n, tmt) %>%
  summarise(mean_osml_est1 = mean(osml_est1),
            mean_osml_est2 = mean(osml_est2),
            mean_osml_est3 = mean(osml_est3))
# exp CEWL
exp_predict_CEWL_means <- exp_predict_CEWL %>%
  dplyr::filter(day_n %in% c(0, 8)) %>%
  group_by(day_n, tmt) %>%
  summarise(mean_CEWL_est = mean(CEWL_est))
# rehydration SMI
rhyd_predict_SMI_means <- rhyd_predict_SMI %>%
  dplyr::filter(day_n == 10) %>%
  group_by(day_n, tmt) %>%
  summarise(mean_SMI_est1 = mean(SMI_est1),
            mean_SMI_est2 = mean(SMI_est2))
# rehydration hct
rhyd_predict_hct_means <- rhyd_predict_hct %>%
  dplyr::filter(day_n == 10) %>%
  group_by(day_n, tmt) %>%
  summarise(mean_hct_est1 = mean(hct_est1),
            mean_hct_est2 = mean(hct_est2),
            mean_hct_est3 = mean(hct_est3)) 
# rehydration osml
rhyd_predict_osml_means <- rhyd_predict_osml %>%
  dplyr::filter(day_n == 10) %>%
  group_by(day_n, tmt) %>%
  summarise(mean_osml_est = mean(osml_est))
```


```{r actual sample means}
actual_SMI_means <- dat %>%
  dplyr::filter(complete.cases(SMI)) %>%
  dplyr::filter(day_n %in% c(0, 8, 10)) %>%
  group_by(tmt, day_n) %>%
  summarise(mean_SMI = mean(SMI))

actual_hct_means <- dat %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  dplyr::filter(day_n %in% c(0, 8, 10)) %>%
  group_by(tmt, day_n) %>%
  summarise(mean_hct = mean(hematocrit_percent))

actual_osml_means <- dat %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  dplyr::filter(day_n %in% c(0, 8, 10)) %>%
  group_by(tmt, day_n) %>%
  summarise(mean_osml = mean(osmolality_mmol_kg_mean))

actual_CEWL_means <- dat %>%
  dplyr::filter(complete.cases(CEWL_g_m2h_mean)) %>%  
  dplyr::filter(day_n %in% c(0, 8, 10)) %>%
  group_by(tmt, day_n) %>%
  summarise(mean_CEWL = mean(CEWL_g_m2h_mean))
```


## SMI 

### Predict

```{r SMI predict fig}
ggplot() + 
  stat_smooth(data = exp_predict_SMI,
              aes(x = day_n,
                  y = SMI_est, 
                  color = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1) + 
  stat_smooth(data = rhyd_predict_SMI,
              aes(x = day_n,
                  y = SMI_est1, 
                  color = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1) + 
  geom_point(data = exp_predict_SMI_means,
             aes(x = day_n, 
                 y = mean_SMI_est, 
                 color = tmt,
                 shape = tmt),
             size = 4) +
  geom_point(data = rhyd_predict_SMI_means,
             aes(x = day_n, 
                 y = mean_SMI_est1, 
                 color = tmt,
                 shape = tmt),
             size = 4) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("Day") + 
  ylab("Body Condition (g)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
) -> SMI_predict_fig
SMI_predict_fig
```



### Actual

```{r SMI actual fig}
ggplot() + 
  geom_line(data = dat,
            aes(x = day_n,
                y = SMI, 
                color = tmt,
                group = individual_ID),
            alpha = 0.2) +
  stat_smooth(data = dat_no_rehab,
              aes(x = day_n,
                  y = SMI, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  stat_smooth(data = dat_for_rehab,
              aes(x = day_n,
                  y = SMI, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1) + 
  geom_point(data = actual_SMI_means,
             aes(x = day_n,
                 y = mean_SMI,
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("Day") + 
  ylab("Body Condition (g)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
) -> SMI_fig
SMI_fig
```

### Actual-Minimal

```{r SMI actual-min fig}
ggplot() + 
  stat_smooth(data = dat_no_rehab,
              aes(x = day_n,
                  y = SMI, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  stat_smooth(data = dat_for_rehab,
              aes(x = day_n,
                  y = SMI, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1) + 
  geom_point(data = actual_SMI_means,
             aes(x = day_n,
                 y = mean_SMI,
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("Day") + 
  ylab("Body Condition (g)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
) -> SMI_fig_min
SMI_fig_min
```



## Hct 

### Predict

```{r hct predict fig}
ggplot() + 
  stat_smooth(data = exp_predict_hct,
              aes(x = day_n,
                  y = hct_est1, 
                  color = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1) + 
  stat_smooth(data = rhyd_predict_hct,
              aes(x = day_n,
                  y = hct_est1, 
                  color = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1) + 
  geom_point(data = exp_predict_hct_means,
             aes(x = day_n, 
                 y = mean_hct_est1, 
                 color = tmt,
                 shape = tmt),
             size = 4) +
  geom_point(data = rhyd_predict_hct_means,
             aes(x = day_n, 
                 y = mean_hct_est1, 
                 color = tmt,
                 shape = tmt),
             size = 4) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Hematocrit (%)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
) -> hct_predict_fig
hct_predict_fig
```




### Actual

```{r hct actual fig}
# need to scrub data
dat_1 <- dat %>%
  dplyr::filter(complete.cases(hematocrit_percent))
dat_for_rehab1 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent))
dat_no_rehab1 <- dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent))

ggplot() +
  geom_line(data = dat_1,
            aes(x = day_n,
                y = hematocrit_percent, 
                color = tmt,
                group = individual_ID),
            alpha = 0.2) +
  stat_smooth(data = dat_no_rehab1,
              aes(x = day_n,
                  y = hematocrit_percent, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  stat_smooth(data = dat_for_rehab1,
              aes(x = day_n,
                  y = hematocrit_percent, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  geom_point(data = actual_hct_means,
             aes(x = day_n,
                 y = mean_hct, 
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Hematocrit (%)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
        ) -> hct_fig
hct_fig
```



### Actual-Minimal

```{r hct actual-minimal fig}
ggplot() +
  stat_smooth(data = dat_no_rehab1,
              aes(x = day_n,
                  y = hematocrit_percent, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  stat_smooth(data = dat_for_rehab1,
              aes(x = day_n,
                  y = hematocrit_percent, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  geom_point(data = actual_hct_means,
             aes(x = day_n,
                 y = mean_hct, 
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Hematocrit (%)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
        ) -> hct_fig_min
hct_fig_min
```



## Osml 

### Predict

```{r osml predict fig}
ggplot() + 
  stat_smooth(data = exp_predict_osml,
              aes(x = day_n,
                  y = osml_est1, 
                  color = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1) + 
  stat_smooth(data = rhyd_predict_osml,
              aes(x = day_n,
                  y = osml_est, 
                  color = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1) + 
  geom_point(data = exp_predict_osml_means,
             aes(x = day_n, 
                 y = mean_osml_est1, 
                 color = tmt,
                 shape = tmt),
             size = 4) +
  geom_point(data = rhyd_predict_osml_means,
             aes(x = day_n, 
                 y = mean_osml_est, 
                 color = tmt,
                 shape = tmt),
             size = 4) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Plasma Osmolality (mmol/kg)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
) -> osml_predict_fig
osml_predict_fig
```


### Actual


```{r osmolality actual fig}
# need to scrub data
dat_2 <- dat %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean))
dat_for_rehab2 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean))
dat_no_rehab2 <- dat_no_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean))


ggplot() + 
  geom_line(data = dat_2,
            aes(x = day_n,
                y = osmolality_mmol_kg_mean, 
                color = tmt,
                group = individual_ID),
            alpha = 0.2) +
  stat_smooth(data = dat_no_rehab2,
              aes(x = day_n,
                  y = osmolality_mmol_kg_mean, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  stat_smooth(data = dat_for_rehab2,
              aes(x = day_n,
                  y = osmolality_mmol_kg_mean, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  geom_point(data = actual_osml_means,
             aes(x = day_n,
                 y = mean_osml,
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Osmolality (mmol/kg)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
        ) -> osml_fig
osml_fig
```


### Actual-Minimal


```{r osmolality actual-minimal fig}
ggplot() + 
  stat_smooth(data = dat_no_rehab2,
              aes(x = day_n,
                  y = osmolality_mmol_kg_mean-2.5, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  stat_smooth(data = dat_for_rehab2,
              aes(x = day_n,
                  y = osmolality_mmol_kg_mean, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  geom_point(data = actual_osml_means,
             aes(x = day_n,
                 y = mean_osml,
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Osmolality (mmol/kg)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
        ) -> osml_fig_min
osml_fig_min
```




## CEWL

### Predict

```{r CEWL predict fig}
ggplot() + 
  stat_smooth(data = exp_predict_CEWL,
              aes(x = day_n,
                  y = CEWL_est, 
                  color = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1) +
  geom_point(data = exp_predict_CEWL_means,
             aes(x = day_n, 
                 y = mean_CEWL_est, 
                 color = tmt,
                 shape = tmt),
             size = 4) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
) -> CEWL_predict_fig
CEWL_predict_fig
```


### Actual


```{r CEWL actual fig}
dat_3 <- dat_no_rehab %>%
  dplyr::filter(complete.cases(CEWL_g_m2h_mean))
ggplot() + 
  geom_line(data = dat_3,
            aes(x = day_n,
                y = CEWL_g_m2h_mean, 
                color = tmt,
                group = individual_ID),
            alpha = 0.2) +
  stat_smooth(data = dat_3,
              aes(x = day_n,
                  y = CEWL_g_m2h_mean, 
                  color = tmt,
                  group = tmt),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  geom_point(data = actual_CEWL_means,
             aes(x = day_n,
                 y = mean_CEWL,
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
        #legend.position = c(0.25,0.85)
        ) -> CEWL_fig
CEWL_fig
```





## Multi-Figures

body condition & blood:

```{r arrange figs}
ggarrange(osml_fig, osml_fig_min, osml_predict_fig,
          hct_fig, hct_fig_min, hct_predict_fig,
          SMI_fig, SMI_fig_min, SMI_predict_fig,
          ncol = 3, nrow = 3,
          common.legend = TRUE,
          legend = "bottom"
          ) -> experiment_multi_fig
experiment_multi_fig
# export figure
ggsave(filename = "experiment_multi_fig.jpeg",
       plot = experiment_multi_fig,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 18, height = 14)
```


CEWL:

```{r arrange CEWL figs}
ggarrange(CEWL_fig, CEWL_predict_fig,
          ncol = 2, nrow = 1,
          common.legend = TRUE,
          legend = "bottom"
          ) -> experiment_CEWL_fig
experiment_CEWL_fig
# export figure
ggsave(filename = "experiment_CEWL_fig.jpeg",
       plot = experiment_CEWL_fig,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 10, height = 6)
```





