---
title: "Climate Water Loss Experiment - Treatment Hydration Analysis"
author: "Savannah Weaver"
date: "2021"
output: pdf_document
toc: TRUE
---

# Packages


```{r setup, include = TRUE, message = FALSE}
`%nin%` = Negate(`%in%`)
if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse") # workflow and plots
if (!require("lme4")) install.packages("lme4")
library("lme4") # for LMMs
if (!require("lmerTest")) install.packages("lmerTest")
library("lmerTest") # for p-values
if (!require("ggpubr")) install.packages("ggpubr")
library("ggpubr") # for multi-ggplot figs
if (!require("UsingR")) install.packages("UsingR")
library("UsingR") # simple.eda model assumption checker
if (!require("broom")) install.packages("broom")
library("broom") # lmer model export
if (!require("broom.mixed")) install.packages("broom.mixed")
library("broom.mixed") # lmer model export
if (!require("car")) install.packages("car")
library("car") # VIFs
if (!require("AICcmodavg")) install.packages("AICcmodavg")
library("AICcmodavg") # model selection
```


# Background and Goals

This data was collected June - August by Master's student Savannah Weaver, advisor Dr. Emily Taylor, and research assistants Tess McIntyre and Taylor Van Rossum. Adult male *Sceloporus occidentalis* were caught across the Cal Poly campus and in Poly Canyon. This R file analyzes the effect of experimental climate treatments on their body condition, osmotic balance, and osmoregulation. Please refer to **doi:** for the published scientific journal article and full details.


# Data

## Load

Read-in data that was compiled, formatted, and checked for completeness in 'capture_analysis'. See that file for information related to the variables. Some variables need to be re-formatted because they lose their data class in export/import. Attach treatment assignments and format that data. Also compute scaled mass index using the equation derived in 'capture_analysis'.

```{r load data}
dat <- read.csv("./data/full_exp_data.csv", # main dataset
                na.strings=c("","NA")) %>%
         # reformat individual_ID to make joining easy
  mutate(individual_ID = as.numeric(individual_ID)) %>%
            # join tmt assignment info
  left_join(read.csv("./data/tmt_assignments.csv"),
            by = "individual_ID") %>%
                # remove irrelevant variables
  dplyr::select(-X, -capture_date, 
                -time_captured, -time_processed, -time_c_temp) %>%
          # format date
  mutate(measurement_date = as.Date(measurement_date,
                                    format = "%Y-%m-%d"),
         # set factor classes
         type = as.factor(type),
         day = as.factor(day),
         individual_ID = as.factor(individual_ID), 
         hemolyzed = as.factor(hemolyzed),
         trial_number = as.factor(trial_number),
         tmt = as.factor(paste(temp_tmt, humidity_tmt)),
         temp_tmt = as.factor(temp_tmt),
         humidity_tmt = as.factor(humidity_tmt),
         conclusion = as.factor(conclusion),
         # set numeric classes
         day_n = as.numeric(day_n),
         mass_g = as.numeric(mass_g),
         hematocrit_percent = as.numeric(hematocrit_percent),
         osmolality_mmol_kg_mean = as.numeric(osmolality_mmol_kg_mean),
         CEWL_g_m2h_mean = as.numeric(CEWL_g_m2h_mean),
         cloacal_temp_C = as.numeric(cloacal_temp_C),
         SVL_mm = as.numeric(SVL_mm),
         # compute scaled mass index
         SMI = mass_g * ((67.71/SVL_mm) ^ 2.65)) %>%
  # only use data from lizards who underwent the full experiment
  # this removes 21 obs for the 7 lizards who didn't complete the exp
  dplyr::filter(conclusion == "complete")
summary(dat)
```



## Clean

There are a handful of points that appear erroneous. 

Individual 239 had post-treatment CEWL >60, which is incredibly unusual for our experiment. He was in the process of shedding when we took this CEWL measurement, so we assume that his process of shedding confounded potential treatment effects, and we should remove that point from our data. We still want to use the other measurements for this individual, so we will just set that CEWL measurement as an NA.

```{r fix erroneous CEWL}
which(dat$CEWL_g_m2h_mean > 60)
dat[332, ]
dat[332, "CEWL_g_m2h_mean"]
dat[332, "CEWL_g_m2h_mean"] <- NA
dat[332, "CEWL_g_m2h_mean"]
dat %>%
  dplyr::filter(complete.cases(CEWL_g_m2h_mean)) %>%
  summarise(max(CEWL_g_m2h_mean))
```


10 trial 1 lizards have unreasonably high osmolality measurements on June 24, which we think are due to an osmometer technical error. The values are way too far outside the usual range to be trustworthy, so we will exclude them.

```{r fix erroneous osml}
which(dat$osmolality_mmol_kg_mean > 500)
dat[136:145, ]
dat[136:145, "osmolality_mmol_kg_mean"]
dat[136:145, "osmolality_mmol_kg_mean"] <- c(rep(NA, 10))
dat[136:145, "osmolality_mmol_kg_mean"]
dat[136:145, ]
dat %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  summarise(max(osmolality_mmol_kg_mean))
```




## Format

Rename some factors:

```{r rename factors}
dat$humidity_tmt <- factor(dat$humidity_tmt,
                           levels = c("humid", "dry"),
                           labels = c("Humid", "Dry"))
dat$temp_tmt <- factor(dat$temp_tmt,
                           levels = c("hot", "cool"),
                           labels = c("Hot", "Cool"))
dat$tmt <- factor(dat$tmt,
                           levels = c("hot humid", "hot dry",
                                      "cool humid", "cool dry"),
                           labels = c("Hot Humid", "Hot Dry",
                                      "Cool Humid", "Cool Dry"))
summary(dat)
```


Make sub-dataframes without rehab data / with only rehab-related data:

```{r sub-dfs}
dat_no_rehab <- dat %>%
  dplyr::filter(type == "exp")
dat_for_rehab <- dat %>%
  dplyr::filter(day_n %in% c(8, 10))
```



## Check

Dates:

```{r check dates}
unique(dat$measurement_date)
```

Number of measurements for each lizard:

```{r check n msmts}
dat_no_rehab %>%
  group_by(individual_ID) %>%
  summarise(n = n()) %>%
  arrange(n)
```


Every lizard has 6 experimental measurements: pre-tmt, mid-tmt, post-tmt, and mass checks on each of the 3 days between mid and post-tmt.


Did any of the treatment groups inherently start out with large differences in response variables?


```{r check starting diffs}
dat %>%
  dplyr::filter(day == "capture") %>%
  group_by(tmt) %>%
  summarise(mean(mass_g),
            mean(SMI),
            mean(hematocrit_percent),
            mean(osmolality_mmol_kg_mean),
            mean(CEWL_g_m2h_mean))
```

There are slight differences, but overall the starting values across groups are more or less the same.




# Experiment Figures

## SMI

```{r SMI exp fig}
# calculate means to overlay
mean_SMI <- dat %>%
  #dplyr::filter(day == "post-exp") %>%
  dplyr::filter(complete.cases(SMI)) %>%
  group_by(tmt, day_n) %>%
  summarise(#day = 8,
            mean_SMI = mean(SMI))

# plot!!
dat_no_rehab %>% 
  ggplot(aes(x = day_n,
             y = SMI, 
             color = tmt,
             )) + 
  
  geom_line(aes(group = individual_ID),
            alpha = 0.2) +
  #stat_smooth(formula = y ~ x, 
   #           method = "lm", 
    #          se = F, 
     #         size = 1.2, 
      #        alpha = 1 ) + 
  # this point is more for the legend than for showing means
  geom_point(data = mean_SMI,
             aes(x = day_n,
                 y = mean_SMI - 0.1, # slight adjustment to center on line
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  geom_line(data = mean_SMI,
             aes(x = day_n,
                 y = mean_SMI,
                 color = tmt)) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Body Condition (g)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
) -> tmt_effects_SMI
tmt_effects_SMI

# export figure
#ggsave(filename = "tmt_effects_SMI.jpeg",
 #      plot = tmt_effects_SMI,
  #     path = "./results_figures",
   #    device = "jpeg",
    #   dpi = 1200,
     #  width = 6, height = 4)
```





## Hct

```{r hct exp fig}
# calculate means to overlay
mean_hct <- dat %>%
  #dplyr::filter(day == "post-exp") %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  group_by(tmt, day_n) %>%
  summarise(#day = 8,
            mean_hct = mean(hematocrit_percent))
# plot
dat_no_rehab %>% 
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  ggplot(aes(x = day_n,
             y = hematocrit_percent, 
             color = tmt,
             )) +
  geom_line(aes(group = individual_ID),
            alpha = 0.2) +
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  # this point is more for the legend than for showing means
  geom_point(data = mean_hct,
             aes(x = day,
                 y = mean_hct, 
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Hematocrit (%)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
        ) -> tmt_effects_hct
tmt_effects_hct

# export figure
#ggsave(filename = "tmt_effects_hct.jpeg",
 #      plot = tmt_effects_hct,
  #     path = "./results_figures",
   #    device = "jpeg",
    #   dpi = 1200,
     #  width = 6, height = 4)
```





## Osmolality


```{r osmolality exp fig}
# calculate means to overlay
mean_osml <- dat %>%
  #dplyr::filter(day == "post-exp") %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  group_by(tmt, day_n) %>%
  summarise(#day = 8,
            mean_osml = mean(osmolality_mmol_kg_mean))
# plot
dat_no_rehab %>% 
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>% 
  ggplot(aes(x = day_n,
             y = osmolality_mmol_kg_mean, 
             color = tmt,
             )) + 
  geom_line(aes(group = individual_ID),
            alpha = 0.2) +
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  # this point is more for the legend than for showing means
  geom_point(data = mean_osml,
             aes(x = day,
                 y = mean_osml + 3, # slight adjust to center on line 
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("Day") + 
  ylab("Osmolality (mmol / kg)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "bottom"
        ) -> tmt_effects_osml
tmt_effects_osml

# export figure
#ggsave(filename = "tmt_effects_osml.jpeg",
 #      plot = tmt_effects_osml,
  #     path = "./results_figures",
   #    device = "jpeg",
    #   dpi = 1200,
     #  width = 6, height = 4)
```



## Multi-Figure

```{r arrange exp figs}
ggarrange(tmt_effects_SMI,
          tmt_effects_hct,
          tmt_effects_osml,
          ncol = 1, nrow = 3,
          common.legend = TRUE,
          legend = "bottom"
          ) -> tmt_multi_fig
tmt_multi_fig
# export figure
#ggsave(filename = "tmt_multi_fig.jpeg",
 #      plot = tmt_multi_fig,
  #     path = "./results_figures",
   #    device = "jpeg",
    #   dpi = 1200,
     #  width = 6, height = 12)
```



## CEWL

```{r CEWL fig}
mean_CEWL <- dat %>%
  dplyr::filter(day == "post-exp") %>%
  dplyr::filter(complete.cases(CEWL_g_m2h_mean)) %>%
  group_by(tmt) %>%
  summarise(day = 8,
            mean_CEWL = mean(CEWL_g_m2h_mean))
dat %>% 
  dplyr::filter(complete.cases(CEWL_g_m2h_mean)) %>%
  ggplot(aes(x = day_n,
             y = CEWL_g_m2h_mean, 
             color = tmt,
             )) + 
  geom_line(aes(group = individual_ID),
            alpha = 0.2) +
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1) + 
  geom_point(data = mean_CEWL,
             aes(x = day,
                 y = mean_CEWL, 
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18),
                     name = "") +
  scale_color_brewer(palette = "Set2",
                     name = "") +
  scale_x_continuous(breaks = c(0, 8),
                     labels = c("0" = "before", "8" = "after")
                     ) +
  xlab("") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 20),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14,
                                 angle = 90),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = c(0.25,0.85)
) -> tmt_effects_CEWL
tmt_effects_CEWL

# export figure
ggsave(filename = "tmt_effects_CEWL.jpeg",
       plot = tmt_effects_CEWL,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 5, height = 4)
```






# Experiment Models

## SMI

### Building

```{r SMI exp model 1}
SMI_mod1 <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                         day_n*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(SMI_mod1)
drop1(SMI_mod1)
car::vif(SMI_mod1) # check collinearity 
```

The full model is the best model based on AIC, but we're going to make all the possible reduced models to compare and determine which is best based on AICc.

drop triple interaction:

```{r SMI exp model 2}
SMI_mod2 <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                         day_n*humidity_tmt*temp_tmt -
                         day_n:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
#summary(SMI_mod2)
car::vif(SMI_mod2) # check collinearity 
```

get only 2 interactions:

```{r SMI exp model 2a}
SMI_mod2a <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                         day_n* (humidity_tmt + temp_tmt) +
                         (1|trial_number/individual_ID))
#summary(SMI_mod2a)
car::vif(SMI_mod2a) # check collinearity 
```

change which variables interact:

```{r SMI exp model 2b}
SMI_mod2b <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                         (day_n + humidity_tmt) * temp_tmt +
                         (1|trial_number/individual_ID))
#summary(SMI_mod2b)
car::vif(SMI_mod2b) # check collinearity 
```

change which variables interact:

```{r SMI exp model 2c}
SMI_mod2c <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                        humidity_tmt* (day_n + temp_tmt) +
                         (1|trial_number/individual_ID))
#summary(SMI_mod2c)
car::vif(SMI_mod2c) # check collinearity 
```

drop to only one interaction:

```{r SMI exp model 3a}
SMI_mod3a <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                         day_n * humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
#summary(SMI_mod3a)
car::vif(SMI_mod3a) # check collinearity 
```

switch which interaction used:

```{r SMI exp model 3b}
SMI_mod3b <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                         day_n * temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
#summary(SMI_mod3b)
car::vif(SMI_mod3b) # check collinearity 
```

switch which interaction used:

```{r SMI exp model 3c}
SMI_mod3c <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                        day_n + humidity_tmt * temp_tmt +
                         (1|trial_number/individual_ID))
#summary(SMI_mod3c)
car::vif(SMI_mod3c) # check collinearity 
```

No interactions:

```{r SMI exp model 4}
SMI_mod4 <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                        humidity_tmt + day_n + temp_tmt +
                         (1|trial_number/individual_ID))
car::vif(SMI_mod4) # check collinearity 
```

drop the temperature effect:

```{r SMI exp model 4a}
SMI_mod4a <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ day_n + humidity_tmt +
                         (1|trial_number/individual_ID))
```

drop the humidity effect:

```{r SMI exp model 4b}
SMI_mod4b <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ day_n + temp_tmt +
                         (1|trial_number/individual_ID))
```

try only humidity and temp:

```{r SMI exp model 4c}
SMI_mod4c <- lme4::lmer(data = dat_no_rehab,
                       SMI ~ 
                        humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
```

test each single effect model:
                   
```{r SMI exp model 5s}
SMI_mod5a <- lme4::lmer(data = dat_no_rehab, SMI ~ day_n +
                         (1|trial_number/individual_ID))
SMI_mod5b <- lme4::lmer(data = dat_no_rehab, SMI ~ temp_tmt +
                         (1|trial_number/individual_ID))
SMI_mod5c <- lme4::lmer(data = dat_no_rehab, SMI ~ humidity_tmt + 
                         (1|trial_number/individual_ID))
```

Finally, null model:

```{r SMI exp model NULL}
SMI_mod_null <- lme4::lmer(data = dat_no_rehab, SMI ~ 1 +
                         (1|trial_number/individual_ID))
```


### Selection

Compare all models:

```{r SMI mod compare}
SMI_models <- list(SMI_mod1, 
                   SMI_mod2, SMI_mod2a, SMI_mod2b, SMI_mod2c, 
                             SMI_mod3a, SMI_mod3b, SMI_mod3c, 
                   SMI_mod4, SMI_mod4a, SMI_mod4b, SMI_mod4c, 
                             SMI_mod5a, SMI_mod5b, SMI_mod5c, 
                   SMI_mod_null)

#specify model names
EXP_mod_names <- c('(model 1) ~ Day, Humd, Temp, Day:Humd, Day:Temp, Humd:Temp, Day:Humd:Temp',
                   '(model 2) ~ Day, Humd, Temp, Day:Humd, Day:Temp, Humd:Temp',
                   '(model 2a) ~ Day, Humd, Temp, Day:Humd, Day:Temp',
                   '(model 2b) ~ Day, Humd, Temp, Day:Temp, Humd:Temp',
                   '(model 2c) ~ Day, Humd, Temp, Day:Humd, Humd:Temp',
                   '(model 3a) ~ Day, Humd, Temp, Day:Humd',
                   '(model 3b) ~ Day, Humd, Temp, Day:Temp',
                   '(model 3c) ~ Day, Humd, Temp, Humd:Temp',
                   '(model 4) ~ Day, Humd, Temp',
                   '(model 4a) ~ Day, Humd',
                   '(model 4b) ~ Day, Temp',
                   '(model 4c) ~ Humd, Temp',
                   '(model 5a) ~ Day',
                   '(model 5b) ~ Temp',
                   '(model 5c) ~ Humd',
                   'null model')

#calculate AIC of each model
SMI_AICc <- data.frame(aictab(cand.set = SMI_models, 
                                 modnames = EXP_mod_names))
SMI_AICc
```


Model 2a is the single best model.


### Assumptions

Check linear regression assumptions/conditions:

```{r SMI exp model assumptions}
plot(SMI_mod2a)
simple.eda(residuals(SMI_mod2a))
shapiro.test(residuals(SMI_mod2a))
```

Normality is very violated, but equal error variance and linearity look fantastic, so I think it's okay.

### Export

Save best model stats with p-values:

```{r best SMI exp model}
SMI_mod_best <- lmerTest::lmer(data = dat_no_rehab, SMI ~ 
                                 day_n* (humidity_tmt + temp_tmt) +
                                 (1|trial_number/individual_ID))
summary(SMI_mod_best)
write.csv(broom.mixed::tidy(SMI_mod_best), 
          "./results_statistics/exp_SMI_best_mod.csv")
write.csv(SMI_AICc, "./results_statistics/exp_SMI_mod_rankings.csv")
```



## Hematocrit

### Building

```{r hct exp model 1}
hct_mod1 <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         day_n*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_mod1)
drop1(hct_mod1)
car::vif(hct_mod1) # check collinearity 
```


drop triple interaction:

```{r hct exp model 2}
hct_mod2 <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         day_n*humidity_tmt*temp_tmt -
                         day_n:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
#summary(hct_mod2)
car::vif(hct_mod2) # check collinearity 
```

get only 2 interactions:

```{r hct exp model 2a}
hct_mod2a <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         day_n* (humidity_tmt + temp_tmt) +
                         (1|trial_number/individual_ID))
#summary(hct_mod2a)
car::vif(hct_mod2a) # check collinearity 
```

change which variables interact:

```{r hct exp model 2b}
hct_mod2b <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         (day_n + humidity_tmt) * temp_tmt +
                         (1|trial_number/individual_ID))
#summary(hct_mod2b)
car::vif(hct_mod2b) # check collinearity 
```

change which variables interact:

```{r hct exp model 2c}
hct_mod2c <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                        humidity_tmt* (day_n + temp_tmt) +
                         (1|trial_number/individual_ID))
#summary(hct_mod2c)
car::vif(hct_mod2c) # check collinearity 
```

drop to only one interaction:

```{r hct exp model 3a}
hct_mod3a <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         day_n * humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
#summary(hct_mod3a)
car::vif(hct_mod3a) # check collinearity 
```

switch which interaction used:

```{r hct exp model 3b}
hct_mod3b <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                         day_n * temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
#summary(hct_mod3b)
car::vif(hct_mod3b) # check collinearity 
```

switch which interaction used:

```{r hct exp model 3c}
hct_mod3c <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                        day_n + humidity_tmt * temp_tmt +
                         (1|trial_number/individual_ID))
#summary(hct_mod3c)
car::vif(hct_mod3c) # check collinearity 
```

No interactions:

```{r hct exp model 4}
hct_mod4 <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                        humidity_tmt + day_n + temp_tmt +
                         (1|trial_number/individual_ID))
car::vif(hct_mod4) # check collinearity 
```

drop the temperature effect:

```{r hct exp model 4a}
hct_mod4a <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ day_n + humidity_tmt +
                         (1|trial_number/individual_ID))
```

drop the humidity effect:

```{r hct exp model 4b}
hct_mod4b <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ day_n + temp_tmt +
                         (1|trial_number/individual_ID))
```

try only humidity and temp:

```{r hct exp model 4c}
hct_mod4c <- lme4::lmer(data = dat_no_rehab,
                       hematocrit_percent ~ 
                        humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
```

test each single effect model:
                   
```{r hct exp model 5s}
hct_mod5a <- lme4::lmer(data = dat_no_rehab, hematocrit_percent ~ day_n +
                         (1|trial_number/individual_ID))
hct_mod5b <- lme4::lmer(data = dat_no_rehab, hematocrit_percent ~ temp_tmt +
                         (1|trial_number/individual_ID))
hct_mod5c <- lme4::lmer(data = dat_no_rehab, hematocrit_percent ~ humidity_tmt + 
                         (1|trial_number/individual_ID))
```

Finally, null model:

```{r hct exp model NULL}
hct_mod_null <- lme4::lmer(data = dat_no_rehab, hematocrit_percent ~ 1 +
                         (1|trial_number/individual_ID))
```


### Selection

Compare all models:

```{r hct mod compare}
hct_models <- list(hct_mod1, 
                   hct_mod2, hct_mod2a, hct_mod2b, hct_mod2c, 
                             hct_mod3a, hct_mod3b, hct_mod3c, 
                   hct_mod4, hct_mod4a, hct_mod4b, hct_mod4c, 
                             hct_mod5a, hct_mod5b, hct_mod5c, 
                   hct_mod_null)

#calculate AIC of each model
hct_AICc <- data.frame(aictab(cand.set = hct_models, 
                                 modnames = EXP_mod_names))
hct_AICc
```


Models 2b, 3b, and 3c all tied for best model.


### Assumptions

Check linear regression assumptions/conditions:

```{r hct exp model assumptions}
plot(hct_mod2b)
simple.eda(residuals(hct_mod2b))
shapiro.test(residuals(hct_mod2b))

plot(hct_mod3b)
simple.eda(residuals(hct_mod3b))
shapiro.test(residuals(hct_mod3b))

plot(hct_mod3c)
simple.eda(residuals(hct_mod3c))
shapiro.test(residuals(hct_mod3c))
```

All assumptions are satisfied (and beautiful). 



### Export

Save best model stats (2b, 3b, 3c) with p-values:

```{r best hct exp models}
hct_mod_best1 <- lmerTest::lmer(data = dat_no_rehab,
                        hematocrit_percent ~ 
                         (day_n + humidity_tmt) * temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_mod_best1)
write.csv(broom.mixed::tidy(hct_mod_best1), 
          "./results_statistics/exp_hct_best_mod1.csv")
hct_mod_best2 <- lmerTest::lmer(data = dat_no_rehab,
                        hematocrit_percent ~ 
                         day_n * temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
summary(hct_mod_best2)
write.csv(broom.mixed::tidy(hct_mod_best2), 
          "./results_statistics/exp_hct_best_mod2.csv")
hct_mod_best3 <- lmerTest::lmer(data = dat_no_rehab,
                        hematocrit_percent ~ 
                         day_n + humidity_tmt * temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_mod_best3)
write.csv(broom.mixed::tidy(hct_mod_best3), 
          "./results_statistics/exp_hct_best_mod3.csv")
write.csv(hct_AICc, "./results_statistics/exp_hct_mod_rankings.csv")
```







## Osmolality


### Building

```{r osml exp model 1}
osml_mod1 <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         day_n*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_mod1)
drop1(osml_mod1)
car::vif(osml_mod1) # check collinearity 
```


drop triple interaction:

```{r osml exp model 2}
osml_mod2 <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         day_n*humidity_tmt*temp_tmt -
                         day_n:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
#summary(osml_mod2)
car::vif(osml_mod2) # check collinearity 
```

get only 2 interactions:

```{r osml exp model 2a}
osml_mod2a <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         day_n* (humidity_tmt + temp_tmt) +
                         (1|trial_number/individual_ID))
#summary(osml_mod2a)
car::vif(osml_mod2a) # check collinearity 
```

change which variables interact:

```{r osml exp model 2b}
osml_mod2b <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         (day_n + humidity_tmt) * temp_tmt +
                         (1|trial_number/individual_ID))
#summary(osml_mod2b)
car::vif(osml_mod2b) # check collinearity 
```

change which variables interact:

```{r osml exp model 2c}
osml_mod2c <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                        humidity_tmt* (day_n + temp_tmt) +
                         (1|trial_number/individual_ID))
#summary(osml_mod2c)
car::vif(osml_mod2c) # check collinearity 
```

drop to only one interaction:

```{r osml exp model 3a}
osml_mod3a <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         day_n * humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
#summary(osml_mod3a)
car::vif(osml_mod3a) # check collinearity 
```

switch which interaction used:

```{r osml exp model 3b}
osml_mod3b <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                         day_n * temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
#summary(osml_mod3b)
car::vif(osml_mod3b) # check collinearity 
```

switch which interaction used:

```{r osml exp model 3c}
osml_mod3c <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                        day_n + humidity_tmt * temp_tmt +
                         (1|trial_number/individual_ID))
#summary(osml_mod3c)
car::vif(osml_mod3c) # check collinearity 
```

No interactions:

```{r osml exp model 4}
osml_mod4 <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                        humidity_tmt + day_n + temp_tmt +
                         (1|trial_number/individual_ID))
car::vif(osml_mod4) # check collinearity 
```

drop the temperature effect:

```{r osml exp model 4a}
osml_mod4a <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ day_n + humidity_tmt +
                         (1|trial_number/individual_ID))
```

drop the humidity effect:

```{r osml exp model 4b}
osml_mod4b <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ day_n + temp_tmt +
                         (1|trial_number/individual_ID))
```

try only humidity and temp:

```{r osml exp model 4c}
osml_mod4c <- lme4::lmer(data = dat_no_rehab,
                       osmolality_mmol_kg_mean ~ 
                        humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
```

test each single effect model:
                   
```{r osml exp model 5s}
osml_mod5a <- lme4::lmer(data = dat_no_rehab, osmolality_mmol_kg_mean ~ day_n +
                         (1|trial_number/individual_ID))
osml_mod5b <- lme4::lmer(data = dat_no_rehab, osmolality_mmol_kg_mean ~ temp_tmt +
                         (1|trial_number/individual_ID))
osml_mod5c <- lme4::lmer(data = dat_no_rehab, osmolality_mmol_kg_mean ~ humidity_tmt + 
                         (1|trial_number/individual_ID))
```

Finally, null model:

```{r osml exp model NULL}
osml_mod_null <- lme4::lmer(data = dat_no_rehab, 
                            osmolality_mmol_kg_mean ~ 1 +
                         (1|trial_number/individual_ID))
```


### Selection

Compare all models:

```{r osml mod compare}
osml_models <- list(osml_mod1, 
                   osml_mod2, osml_mod2a, osml_mod2b, osml_mod2c, 
                             osml_mod3a, osml_mod3b, osml_mod3c, 
                   osml_mod4, osml_mod4a, osml_mod4b, osml_mod4c, 
                             osml_mod5a, osml_mod5b, osml_mod5c, 
                   osml_mod_null)

#calculate AIC of each model
osml_AICc <- data.frame(aictab(cand.set = osml_models, 
                                 modnames = EXP_mod_names))
osml_AICc
```


Models 1, 2, and 2b, are all tied for best model.


### Assumptions

Check linear regression assumptions/conditions:

```{r osml exp model assumptions}
plot(osml_mod1)
simple.eda(residuals(osml_mod1))
shapiro.test(residuals(osml_mod1))

plot(osml_mod2)
simple.eda(residuals(osml_mod2))
shapiro.test(residuals(osml_mod2))

plot(osml_mod2b)
simple.eda(residuals(osml_mod2b))
shapiro.test(residuals(osml_mod2b))
```

All assumptions are OKAY.



### Export

Save best model stats (1, 2, 2b) with p-values:

```{r best osml exp models}
osml_mod_best1 <- lmerTest::lmer(data = dat_no_rehab,
                        osmolality_mmol_kg_mean ~ 
                         day_n * humidity_tmt * temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_mod_best1)
write.csv(broom.mixed::tidy(osml_mod_best1), 
          "./results_statistics/exp_osml_best_mod1.csv")
osml_mod_best2 <- lmerTest::lmer(data = dat_no_rehab,
                        osmolality_mmol_kg_mean ~ 
                         day_n * humidity_tmt * temp_tmt -
                          day_n:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_mod_best2)
write.csv(broom.mixed::tidy(osml_mod_best2), 
          "./results_statistics/exp_osml_best_mod2.csv")
osml_mod_best3 <- lmerTest::lmer(data = dat_no_rehab,
                        osmolality_mmol_kg_mean ~ 
                         (day_n + humidity_tmt) * temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_mod_best3)
write.csv(broom.mixed::tidy(osml_mod_best3), 
          "./results_statistics/exp_osml_best_mod3.csv")
write.csv(osml_AICc, "./results_statistics/exp_osml_mod_rankings.csv")
```




## CEWL


### Building



```{r CEWL exp model 1}
dat_no_rehab_rm <- dat_no_rehab %>%
  dplyr::filter(complete.cases(.))
CEWL_mod1 <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C + 
                         day_n*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(CEWL_mod1)
drop1(CEWL_mod1)
car::vif(CEWL_mod1) # check collinearity 
```


drop triple interaction:

```{r CEWL exp model 2}
CEWL_mod2 <- lme4::lmer(data = dat_no_rehab_rm,
                       CEWL_g_m2h_mean ~ cloacal_temp_C + 
                         day_n*humidity_tmt*temp_tmt -
                         day_n:humidity_tmt:temp_tmt +
                         (1|trial_number/individual_ID))
summary(CEWL_mod2)
drop1(CEWL_mod2)
car::vif(CEWL_mod2) # check collinearity 
```



get only 2 interactions:

```{r CEWL exp model 2a}
CEWL_mod2a <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ cloacal_temp_C + 
                         day_n* (humidity_tmt + temp_tmt) +
                         (1|trial_number/individual_ID))
#summary(CEWL_mod2a)
drop1(CEWL_mod2a)
car::vif(CEWL_mod2a) # check collinearity 
```

change which variables interact:

```{r CEWL exp model 2b}
CEWL_mod2b <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ cloacal_temp_C + 
                         (day_n + humidity_tmt) * temp_tmt +
                         (1|trial_number/individual_ID))
#summary(CEWL_mod2b)
drop1(CEWL_mod2b)
car::vif(CEWL_mod2b) # check collinearity 
```

change which variables interact:

```{r CEWL exp model 2c}
CEWL_mod2c <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ cloacal_temp_C + 
                        humidity_tmt* (day_n + temp_tmt) +
                         (1|trial_number/individual_ID))
#summary(CEWL_mod2c)
drop1(CEWL_mod2c)
car::vif(CEWL_mod2c) # check collinearity 
```


test dropping cloacal temperature from each:

```{r CEWL exp model 2abc no cloacal temp}
CEWL_mod2ax <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ 
                         day_n* (humidity_tmt + temp_tmt) +
                         (1|trial_number/individual_ID))
CEWL_mod2bx <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ 
                         (day_n + humidity_tmt) * temp_tmt +
                         (1|trial_number/individual_ID))
CEWL_mod2cx <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ 
                        humidity_tmt* (day_n + temp_tmt) +
                         (1|trial_number/individual_ID))
```

drop to only one interaction:

```{r CEWL exp model 3a}
CEWL_mod3a <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ cloacal_temp_C + 
                         day_n * humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
#summary(CEWL_mod3a)
car::vif(CEWL_mod3a) # check collinearity 
```

switch which interaction used:

```{r CEWL exp model 3b}
CEWL_mod3b <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ cloacal_temp_C + 
                         day_n * temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
#summary(CEWL_mod3b)
car::vif(CEWL_mod3b) # check collinearity 
```

switch which interaction used:

```{r CEWL exp model 3c}
CEWL_mod3c <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ cloacal_temp_C + 
                        day_n + humidity_tmt * temp_tmt +
                         (1|trial_number/individual_ID))
#summary(CEWL_mod3c)
car::vif(CEWL_mod3c) # check collinearity 
```

test dropping cloacal temperature from each:

```{r CEWL exp model 3abc without cloacal temp}
CEWL_mod3ax <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ 
                         day_n * humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
CEWL_mod3bx <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ 
                         day_n * temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
CEWL_mod3cx <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ 
                        day_n + humidity_tmt * temp_tmt +
                         (1|trial_number/individual_ID))
```


No interactions:

```{r CEWL exp model 4}
CEWL_mod4 <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ 
                        humidity_tmt + day_n + temp_tmt +
                         (1|trial_number/individual_ID))
car::vif(CEWL_mod4) # check collinearity 
```

drop the temperature effect:

```{r CEWL exp model 4a}
CEWL_mod4a <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ 
                         day_n + humidity_tmt +
                         (1|trial_number/individual_ID))
```

drop the humidity effect:

```{r CEWL exp model 4b}
CEWL_mod4b <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ 
                         day_n + temp_tmt +
                         (1|trial_number/individual_ID))
```

try only humidity and temp:

```{r CEWL exp model 4c}
CEWL_mod4c <- lme4::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ 
                        humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
```

test each single effect model:
                   
```{r CEWL exp model 5s}
CEWL_mod5a <- lme4::lmer(data = dat_no_rehab, CEWL_g_m2h_mean ~ 
                           day_n +
                         (1|trial_number/individual_ID))
CEWL_mod5b <- lme4::lmer(data = dat_no_rehab, CEWL_g_m2h_mean ~ 
                           temp_tmt +
                         (1|trial_number/individual_ID))
CEWL_mod5c <- lme4::lmer(data = dat_no_rehab, CEWL_g_m2h_mean ~
                           humidity_tmt + 
                         (1|trial_number/individual_ID))
```

Finally, null model:

```{r CEWL exp model NULL}
CEWL_mod_null <- lme4::lmer(data = dat_no_rehab, CEWL_g_m2h_mean ~ 1 +
                         (1|trial_number/individual_ID))
```


### Selection

Compare all models:

```{r CEWL mod compare}
CEWL_models <- list(CEWL_mod1, 
                   CEWL_mod2, CEWL_mod2a, CEWL_mod2b, CEWL_mod2c, 
                              CEWL_mod2ax,CEWL_mod2bx,CEWL_mod2cx,
                              CEWL_mod3a, CEWL_mod3b, CEWL_mod3c, 
                              CEWL_mod3ax,CEWL_mod3bx,CEWL_mod3cx,
                   CEWL_mod4, CEWL_mod4a, CEWL_mod4b, CEWL_mod4c, 
                              CEWL_mod5a, CEWL_mod5b, CEWL_mod5c, 
                   CEWL_mod_null)
CEWL_mod_names <- c('(model 1) ~ Temp-Clo, Day, Humd, Temp, Day:Humd, Day:Temp, Humd:Temp, Day:Humd:Temp',
                   '(model 2) ~ Temp-Clo, Day, Humd, Temp, Day:Humd, Day:Temp, Humd:Temp',
                   '(model 2a) ~ Temp-Clo, Day, Humd, Temp, Day:Humd, Day:Temp',
                   '(model 2b) ~ Temp-Clo, Day, Humd, Temp, Day:Temp, Humd:Temp',
                   '(model 2c) ~ Temp-Clo, Day, Humd, Temp, Day:Humd, Humd:Temp',
                   '(model 2ax) ~ Day, Humd, Temp, Day:Humd, Day:Temp',
                   '(model 2bx) ~ Day, Humd, Temp, Day:Temp, Humd:Temp',
                   '(model 2cx) ~ Day, Humd, Temp, Day:Humd, Humd:Temp',
                   '(model 3a) ~ Temp-Clo, Day, Humd, Temp, Day:Humd',
                   '(model 3b) ~ Temp-Clo, Day, Humd, Temp, Day:Temp',
                   '(model 3c) ~ Temp-Clo, Day, Humd, Temp, Humd:Temp',
                   '(model 3ax) ~ Day, Humd, Temp, Day:Humd',
                   '(model 3bx) ~ Day, Humd, Temp, Day:Temp',
                   '(model 3cx) ~ Day, Humd, Temp, Humd:Temp',
                   '(model 4) ~ Day, Humd, Temp',
                   '(model 4a) ~ Day, Humd',
                   '(model 4b) ~ Day, Temp',
                   '(model 4c) ~ Humd, Temp',
                   '(model 5a) ~ Day',
                   '(model 5b) ~ Temp',
                   '(model 5c) ~ Humd',
                   'null model')

#calculate AIC of each model
CEWL_AICc <- data.frame(aictab(cand.set = CEWL_models, 
                                 modnames = CEWL_mod_names))
CEWL_AICc
```


the complete, full model is the best model.


### Assumptions

Check linear regression assumptions/conditions:

```{r CEWL exp model assumptions}
plot(CEWL_mod1)
simple.eda(residuals(CEWL_mod1))
shapiro.test(residuals(CEWL_mod1))
```

normality is not good, both other assumptions OKAY enough.



### Export

Save best model stats with p-values:

```{r best CEWL exp model}
CEWL_mod_best <- lmerTest::lmer(data = dat_no_rehab,
                       CEWL_g_m2h_mean ~ cloacal_temp_C + 
                         day_n*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(CEWL_mod_best)
write.csv(broom.mixed::tidy(CEWL_mod_best), 
          "./results_statistics/exp_CEWL_best_mod.csv")
write.csv(CEWL_AICc, "./results_statistics/exp_CEWL_mod_rankings.csv")
```





# Rehydration Figures


## SMI

```{r SMI rehydration fig}
# calculate means to overlay
mean_SMI_rehab <- dat %>%
  dplyr::filter(type == "rehab") %>%
  dplyr::filter(complete.cases(SMI)) %>%
  group_by(tmt) %>%
  summarise(day = 10,
            mean_SMI = mean(SMI))

# plot!!
dat_for_rehab %>% 
  ggplot(aes(x = day_n,
             y = SMI, 
             color = tmt,
             )) + 
  geom_line(aes(group = individual_ID),
            alpha = 0.2) +
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  # this point is more for the legend than for showing means
  geom_point(data = mean_SMI_rehab,
             aes(x = day,
                 y = mean_SMI,
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Body Condition (g)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
) -> rehab_effects_SMI
rehab_effects_SMI

# export figure
ggsave(filename = "rehab_effects_SMI.jpeg",
       plot = rehab_effects_SMI,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 6, height = 4)
```




## Hematocrit

```{r hct rehydration fig}
# calculate means to overlay
mean_hct_rehab <- dat %>%
  dplyr::filter(type == "rehab") %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  group_by(tmt) %>%
  summarise(day = 10,
            mean_hct = mean(hematocrit_percent))
# plot
dat_for_rehab %>% 
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  ggplot(aes(x = day_n,
             y = hematocrit_percent, 
             color = tmt,
             )) +
  geom_line(aes(group = individual_ID),
            alpha = 0.2) +
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  # this point is more for the legend than for showing means
  geom_point(data = mean_hct_rehab,
             aes(x = day,
                 y = mean_hct, 
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Hematocrit (%)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
        ) -> rehab_effects_hct
rehab_effects_hct

# export figure
ggsave(filename = "rehab_effects_hct.jpeg",
       plot = rehab_effects_hct,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 6, height = 4)
```





## Osmolality

```{r osmolality rehydration fig}
# calculate means to overlay
mean_osml_rehab <- dat %>%
  dplyr::filter(type == "rehab") %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  group_by(tmt) %>%
  summarise(day = 10,
            mean_osml = mean(osmolality_mmol_kg_mean))
# plot
dat_for_rehab %>% 
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>% 
  ggplot(aes(x = day_n,
             y = osmolality_mmol_kg_mean, 
             color = tmt,
             )) + 
  geom_line(aes(group = individual_ID),
            alpha = 0.2) +
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  # this point is more for the legend than for showing means
  geom_point(data = mean_osml_rehab,
             aes(x = day,
                 y = mean_osml,
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("Day") + 
  ylab("Osmolality (mmol / kg)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "bottom"
        ) -> rehab_effects_osml
rehab_effects_osml

# export figure
ggsave(filename = "rehab_effects_osml.jpeg",
       plot = rehab_effects_osml,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 6, height = 4)
```



## Multi-Figure

```{r arrange rehydration figs}
ggarrange(rehab_effects_SMI,
          rehab_effects_hct,
          rehab_effects_osml,
          ncol = 1, nrow = 3,
          common.legend = TRUE,
          legend = "bottom"
          ) -> rehab_multi_fig
rehab_multi_fig
# export figure
ggsave(filename = "rehab_multi_fig.jpeg",
       plot = rehab_multi_fig,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 6, height = 12)
```



# Rehydration Models


## SMI

```{r SMI rehydration model 1}
SMI_rmod1 <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ 
                         type*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(SMI_rmod1)
drop1(SMI_rmod1)
```

Remove three-way interaction and day*humidity interaction.

```{r SMI rehydration model 2}
SMI_rmod2 <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ 
                         type*temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
summary(SMI_rmod2)
drop1(SMI_rmod2)
```

Remove temperature and humidity single effects:

```{r SMI rehydration model 3}
SMI_rmod3 <- lme4::lmer(data = dat_for_rehab,
                       SMI ~ 
                         type:temp_tmt + type +
                         (1|trial_number/individual_ID))
summary(SMI_rmod3)
drop1(SMI_rmod3)
```

This is the best model.

Check linear regression assumptions/conditions:

```{r SMI rehydration model assumptions}
plot(SMI_rmod3)
simple.eda(residuals(SMI_rmod3))
shapiro.test(residuals(SMI_rmod3))
```

Save best model stats:

```{r best SMI rehydration model}
SMI_rmod_best <- lmerTest::lmer(data = dat_for_rehab,
                       SMI ~ 
                         type:temp_tmt + type +
                         (1|trial_number/individual_ID))
summary(SMI_rmod_best)
write.csv(broom.mixed::tidy(SMI_rmod_best), 
          "./results_statistics/rehab_effects_SMI.csv")
```



## Hematocrit

```{r hct rehydration model 1}
hct_rmod1 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
                       hematocrit_percent ~ 
                         type*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_rmod1)
drop1(hct_rmod1)
```

Remove humidity from the three-way interaction:

```{r hct rehydration model 2}
hct_rmod2 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
                       hematocrit_percent ~ 
                         type*temp_tmt + humidity_tmt +
                         (1|trial_number/individual_ID))
summary(hct_rmod2)
drop1(hct_rmod2)
```


Drop humidity from the model completely:

```{r hct rehydration model 3}
hct_rmod3 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
                       hematocrit_percent ~ 
                         type * temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_rmod3)
drop1(hct_rmod3)
```

The model is the same whether we drop the interaction or not, so we choose to drop it to simplify the model.

```{r hct rehydration model 4}
hct_rmod4 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
                       hematocrit_percent ~ 
                         type + temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_rmod4)
drop1(hct_rmod4)
```

This seems to be the best model.

Check linear regression assumptions/conditions:

```{r hct rehydration model assumptions}
plot(hct_rmod4)
simple.eda(residuals(hct_rmod4))
shapiro.test(residuals(hct_rmod4))
```

Save best model stats:

```{r best hct rehydration model}
hct_rmod_best <- dat_for_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lmerTest::lmer(data = .,
                       hematocrit_percent ~ 
                         type + temp_tmt +
                         (1|trial_number/individual_ID))
summary(hct_rmod_best)
write.csv(broom::tidy(hct_rmod_best), 
          "./results_statistics/rehab_effects_hct.csv")
```



## Osmolality

```{r osmolality rehydration model 1}
osml_rmod1 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  lme4::lmer(data = .,
                       osmolality_mmol_kg_mean ~ 
                         type*humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_rmod1)
drop1(osml_rmod1)
```

Drop the interaction with day:

```{r osmolality rehydration model 2}
osml_rmod2 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  lme4::lmer(data = .,
                       osmolality_mmol_kg_mean ~ 
                         type + humidity_tmt*temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_rmod2)
drop1(osml_rmod2)
```


Drop the interaction between temperature and humidity:

```{r osmolality rehydration model 3}
osml_rmod3 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  lme4::lmer(data = .,
                       osmolality_mmol_kg_mean ~ 
                         type + humidity_tmt + temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_rmod3)
drop1(osml_rmod3)
```

Drop humidity:

```{r osmolality rehydration model 4}
osml_rmod4 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  lme4::lmer(data = .,
                       osmolality_mmol_kg_mean ~ 
                         type + temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_rmod4)
drop1(osml_rmod4)
```

The best model to predict the rehydration effects on plasma osmolality included only additive effects of rehab and acclimation temp tmt prior to rehab.

Check linear regression assumptions/conditions:

```{r osml rehydration model assumptions}
plot(osml_rmod4)
simple.eda(residuals(osml_rmod4))
shapiro.test(residuals(osml_rmod4))
```

Save best model stats:

```{r best osmolality rehydration model}
osml_rmod_best <- dat_for_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  lmerTest::lmer(data = .,
                       osmolality_mmol_kg_mean ~ 
                         type + temp_tmt +
                         (1|trial_number/individual_ID))
summary(osml_rmod_best)
write.csv(broom::tidy(osml_rmod_best), 
          "./results_statistics/rehab_effects_osml.csv")
```




## Conclusion

During the 2-day rehydration opportunity provided to lizards following experimental climate treatment, body condition and plasma osmolality significantly increased and hematocrit significantly decreased. The change in body condition during rehydration was also effected by the interaction between rehydration and acclimation temperature prior to rehydration, with cool-acclimated lizards having overall higher body condition and showing a larger increase in body condition during rehydration. Cool-acclimated lizards also maintained higher hematocrit and lower plasma osmolality during rehydration compared to other treatment groups, although changes in hct and osml during rehydration were not different due to acclimation. Nested effects of trial number and individual lizard ID were included as random effects in all three LMMs.


This is true for pretty much all the model selection I did:
In model selection, each dropped variable only improved the model by <2-delta-AIC, so each stepwise model predicted the variability in the response variable equally well. However, the variables dropped also had very low t-values, evidence for their lack of influence thus exclusion from the model.



# Exp -> Rehydration Figures

## Calculate Predicted Values

```{r mod predicts for plots}
dat_no_rehab_predict_CEWL <- dat_no_rehab %>%
  dplyr::filter(complete.cases(CEWL_g_m2h_mean)) %>%
  mutate(CEWL_est = predict(CEWL_mod_best))
dat_no_rehab_predict_hct <- dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  mutate(hct_est1 = predict(hct_mod_best1),
         hct_est2 = predict(hct_mod_best2),
         hct_est3 = predict(hct_mod_best3))
dat_no_rehab_predict_osml <- dat_no_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>%
  mutate(osml_est1 = predict(osml_mod_best1),
         osml_est2 = predict(osml_mod_best2),
         osml_est3 = predict(osml_mod_best3))
```


## SMI

```{r SMI fig}
ggplot() + 
  geom_line(data = dat_no_rehab,
            aes(x = day_n,
             y = SMI, 
             color = tmt,
             group = individual_ID),
            alpha = 0.2) +
  geom_line(data = dat_for_rehab,
            aes(x = day_n,
             y = SMI, 
             color = tmt,
             group = individual_ID),
            alpha = 0.2) +
  stat_smooth(data = dat_no_rehab,
    aes(x = day_n,
             y = SMI, 
             color = tmt,
             group = tmt),
            formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  stat_smooth(data = dat_for_rehab,
    aes(x = day_n,
             y = SMI, 
             color = tmt,
             group = tmt),
            formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  geom_point(data = mean_SMI,
             aes(x = day,
                 y = mean_SMI - 0.1, # slight adjustment to center on line
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  geom_point(data = mean_SMI_rehab,
             aes(x = day,
                 y = mean_SMI,
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("Day") + 
  ylab("Body Condition (g)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
) -> overall_fig_SMI
overall_fig_SMI

# export figure
ggsave(filename = "overall_fig_SMI.jpeg",
       plot = overall_fig_SMI,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 6, height = 4)
```





## Hct

```{r hct fig}
# need to scrub data
dat_for_rehab1 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent))
dat_no_rehab1 <- dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent))

# plot
ggplot() +
  geom_line(data = dat_no_rehab1,
            aes(x = day_n,
             y = hematocrit_percent, 
             color = tmt,
             group = individual_ID),
            alpha = 0.2) +
  geom_line(data = dat_for_rehab1,
            aes(x = day_n,
             y = hematocrit_percent, 
             color = tmt,
             group = individual_ID),
            alpha = 0.2) +
  stat_smooth(data = dat_no_rehab1,
            aes(x = day_n,
             y = hematocrit_percent, 
             color = tmt,
             group = tmt),
            formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  stat_smooth(data = dat_for_rehab1,
            aes(x = day_n,
             y = hematocrit_percent, 
             color = tmt,
             group = tmt),
            formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  geom_point(data = mean_hct,
             aes(x = day,
                 y = mean_hct, 
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  geom_point(data = mean_hct_rehab,
             aes(x = day,
                 y = mean_hct, 
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Hematocrit (%)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "none"
        ) -> overall_fig_hct
overall_fig_hct

# export figure
ggsave(filename = "overall_fig_hct.jpeg",
       plot = overall_fig_hct,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 6, height = 4)
```





## Osmolality


```{r osmolality fig}
# need to scrub data
dat_for_rehab2 <- dat_for_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean))
dat_no_rehab2 <- dat_no_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean))

# plot
ggplot() + 
  geom_line(data = dat_no_rehab2,
            aes(x = day_n,
             y = osmolality_mmol_kg_mean, 
             color = tmt,
             group = individual_ID),
            alpha = 0.2) +
  geom_line(data = dat_for_rehab2,
            aes(x = day_n,
             y = osmolality_mmol_kg_mean, 
             color = tmt,
             group = individual_ID),
            alpha = 0.2) +
  stat_smooth(data = dat_no_rehab2,
            aes(x = day_n,
             y = osmolality_mmol_kg_mean, 
             color = tmt,
             group = tmt),
            formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  stat_smooth(data = dat_for_rehab2,
            aes(x = day_n,
             y = osmolality_mmol_kg_mean, 
             color = tmt,
             group = tmt),
            formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.2, 
              alpha = 1 ) + 
  geom_point(data = mean_osml,
             aes(x = day,
                 y = mean_osml + 3, # slight adjust to center on line 
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  geom_point(data = mean_osml_rehab,
             aes(x = day,
                 y = mean_osml,
                 color = tmt,
                 shape = tmt),
             size = 4, 
             alpha = 1) +
  theme_classic() + 
  scale_shape_manual(values = c(15:18), name = "") +
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8, 10)) +
  scale_color_brewer(palette = "Set2", name = "") +
  xlab("") + 
  ylab("Osmolality (mmol / kg)") + 
  guides(shape = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_vline(xintercept = 9,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 22),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 16),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 22),
        legend.text.align = 0,
        legend.position = "bottom"
        ) -> overall_fig_osml
overall_fig_osml

# export figure
ggsave(filename = "overall_fig_osml.jpeg",
       plot = overall_fig_osml,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 6, height = 4)
```



## Multi-Figure

```{r arrange figs}
ggarrange(overall_fig_osml,
          overall_fig_hct,
          overall_fig_SMI,
          ncol = 1, nrow = 3,
          common.legend = TRUE,
          legend = "bottom"
          ) -> overall_multi_fig
overall_multi_fig
# export figure
ggsave(filename = "overall_multi_fig.jpeg",
       plot = overall_multi_fig,
       path = "./results_figures",
       device = "jpeg",
       dpi = 1200,
       width = 6, height = 12)
```








