---
title: "HOBO Data Logger Climate Treatment Summary"
author: "Savannah Weaver"
date: "10/15/2021"
output: pdf_document
toc: TRUE
---

# Packages

```{r setup, echo=T, results='hide', message=FALSE}
if (!require("tidyverse")) install.packages("tidyverse") 
library("tidyverse")
```

**remember to do rehab data separately**

# Load in Data

This data was collected using Onset HOBO temperature and humidity dataloggers during the course of our experiment. See __ for full details. 

The data is in a separate file for each download for each logger, so I need to compile each of those into one dataset. 

To do this, first I compile a list of the filenames I need to read-in.

## File Names

```{r, filenames}
# make a list of file names of all data to load in
filenames <- list.files(path = "data/HOBOs")
```

Next, I make a function that will read in the data from each csv, name and organize the data correctly. 


## Function

```{r, create read_HOBO_files function to get each csv data}
# make a function to read in data from each csv file and add correct identifiers
read_HOBO_files <- function(filename) {
  
  # edit the filename inputted to funtion
  # to make a unique identifier for each logger
  name <- substr(filename, 1, nchar(filename)-15)
  # read in the csv file for this given filename
  dat <- read.csv(file.path("data/HOBOs", filename),
                # each csv has headers
                header = TRUE,
                # this is what I want to rename the col headers
                col.names = c("order", "date_time_PST", "temp_C", 
                              "relative_humidity", "dew_pt_C", 
                              # the 6,7,8th cols are not data
                              # logger use info we don't need
                              "mostly_blank", "mostly_blank", "mostly_blank")
                ) %>%
    # select only the cols with data we want
    # don't need order- just an arbitrary observation identifier
    # don't need "mostly_blank" cols- unnecessary logger use info
    # but get the rest of the cols with informative data
    dplyr::select(date_time_PST, temp_C, relative_humidity, dew_pt_C) %>%
    # add a column with the name of the HOBO the data is from 
    dplyr::mutate(HOBO_ID = name)
  
  # return the dataframe for that single csv file
  dat
}
```



## Apply

Finally, I apply the function I made to all of the filenames I compiled, then put all of those dataframes into one dataframe for my analyses.

This will print warnings saying that header and col.names are different lengths, because the data has extra notes on logger usage that we read-in, but get rid of.

```{r, apply read_HOBO_files function}
# apply function to get data from all csvs
all_HOBO_data <- lapply(filenames, read_HOBO_files) %>%
  # paste all data files together into one df by row
  reduce(rbind) %>% 
  mutate(date_time_PST = as.POSIXct(date_time_PST, 
                                    format = "%m/%d/%Y %H:%M:%S")) %>%
  # remove missing data
  dplyr::filter(complete.cases(date_time_PST, relative_humidity)) %>%
  mutate(serial = (substr(HOBO_ID, 7, 15)), # select characters
         serial = str_trim(serial), # remove trailing white space
         serial = str_replace_all(serial, "[^A-z0-9]", "_"), # replace any special characters with underscores
         serial = as.factor(serial), # set class as factor
         temp_C = as.numeric(temp_C),
         relative_humidity = as.numeric(relative_humidity),
         dew_pt_C = as.numeric(dew_pt_C)
         )
summary(all_HOBO_data)
head(all_HOBO_data)
```



# Data Wrangling

I need to subset the data based on the date range for actual experiment days.

Trail 1: June 16-24
Trail 2: June 26 - July 4
Trial 3: July 20-28
Trial 4: August 8-16
Trial 5: August 22-30
Trial 5 rehab: August 31-32

However, lizards were usually placed in the chambers late on the first day and taken out early on the last day, so I'll use the exclusive range of those dates (17-23 instead of 16-24). the list of dates should total 7 days each trial * 5 trials = 30 dates.


## Dates

```{r}
t1_days <- seq(as.Date("2021-06-17"),
               as.Date("2021-06-23"), "days")
t2_days <- seq(as.Date("2021-06-27"),
               as.Date("2021-07-03"), "days")
t3_days <- seq(as.Date("2021-07-21"),
               as.Date("2021-07-27"), "days")
t4_days <- seq(as.Date("2021-08-09"),
               as.Date("2021-08-15"), "days")
t5_days <- seq(as.Date("2021-08-23"),
               as.Date("2021-08-29"), "days")
exp_dates <- c(t1_days, t2_days, t3_days, t4_days, t5_days)
```

Subset data:

```{r}
subset_HOBO_data <- all_HOBO_data %>%
  mutate(date_only = as.Date(date_time_PST, format = "%Y-%m-%d")) %>%
  dplyr::filter(date_only %in% exp_dates)
```


## Trial & Tmt

```{r}
# load HOBO assignment data
HOBO_tmts <- read.csv("./data/HOBO_assignments.csv") %>%
  dplyr::filter(trial != "5 rehab") %>%
  mutate(trial = as.factor(trial),
         serial = as.factor(serial),
         humidity_tmt = as.factor(humidity_tmt),
         temp_tmt = as.factor(temp_tmt)
         ) %>%
  dplyr::select(-notes, -in.google.drive.) 
# combine
format_HOBO_data <- subset_HOBO_data %>% 
  mutate(trial = as.factor(case_when(date_only %in% t1_days ~ 1,
                                     date_only %in% t2_days ~ 2,
                                     date_only %in% t3_days ~ 3,
                                     date_only %in% t4_days ~ 4,
                                     date_only %in% t5_days ~ 5))
         ) %>%
  left_join(HOBO_tmts, by = c("trial", "serial"))
summary(format_HOBO_data)
```



# Check Data

## Dates

```{r}
format_HOBO_data %>% 
  group_by(trial) %>%
  summarise(min_date = (min(date_only)),
            max_date = (max(date_only)))
```






Check temperature values:

```{r}
format_HOBO_data %>% 
  ggplot(data = ., aes(x = temp_C)) +
  geom_histogram()
format_HOBO_data %>%
  group_by(date_only, HOBO_ID) %>%
  summarise(mean_temp = mean(temp_C)) %>%
  arrange(mean_temp, date_only)
```


# Data Cleaning

set the not in function:

```{r}
`%nin%` = Negate(`%in%`)
```


## HOBO "dry cool 1":

```{r}
# temperature
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Dry_Cool_1") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = temp_C, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "red") + 
  geom_line(color = "red") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-27")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-29")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-08")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-11")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-12")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
# humidity
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Dry_Cool_1") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = relative_humidity, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "blue") + 
  geom_line(color = "blue") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-27")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-29")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-08")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-11")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-12")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
```

There is room temperature/wacky temperature data between April 27-29, and on May 4, 8, 11, and 12. The humidity data only looks wrong on May 11 & 12. 

Try excluding. For some reason the dates plotted versus filtered line up differently, which is why I filter out the date before what I plotted.

```{r}
unwanted_dates_dc1 <- as.Date(c("2021-04-26", "2021-04-27", 
                        "2021-04-28",
                        "2021-05-03", "2021-05-07", 
                        "2021-05-10", "2021-05-11"),
                        format = "%Y-%m-%d")
dc1_subset <- format_HOBO_data %>%
  dplyr::filter(HOBO_ID == "Dry_Cool_1") %>%
  dplyr::filter(date_only %nin% unwanted_dates_dc1)
```

plot again to show that we removed erroneous data:

```{r}
# temperature
dc1_subset %>% 
  ggplot(data = ., aes(x = date_time_PST, 
                       y = temp_C, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "red") + 
  geom_line(color = "red") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-27")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-29")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-08")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-11")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-12")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
# humdity
dc1_subset %>% 
  ggplot(data = ., aes(x = date_time_PST, 
                       y = relative_humidity, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "blue") + 
  geom_line(color = "blue") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-27")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-29")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-08")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-11")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-12")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
```

And, we confirmed that HOBO "dry cool 1" was indeed in the dry and cool treatment.


## HOBO "dry cool 2":

```{r}
# temperature
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Dry_Cool_2") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = temp_C, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "red") + 
  geom_line(color = "red") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
# humidity
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Dry_Cool_2") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = relative_humidity, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "blue") + 
  geom_line(color = "blue") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-20")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-25")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-27")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-19")), 
             linetype = 4, color = "purple") +
  theme_classic()
```

Temperature is wrong on May 4. The humidity is wrong April 19-20, April 25-27, and on May 19. 

Try excluding. For some reason the dates plotted versus filtered line up differently, which is why I filter out the date before what I plotted.

```{r}
unwanted_dates_dc2 <- as.Date(c("2021-04-18", "2021-04-19", 
                        "2021-04-24", "2021-04-25", "2021-04-26",
                        "2021-05-03", "2021-05-18"),
                        format = "%Y-%m-%d")
dc2_subset <- format_HOBO_data %>%
  dplyr::filter(HOBO_ID == "Dry_Cool_2") %>%
  dplyr::filter(date_only %nin% unwanted_dates_dc2)
```

plot again to show that we removed erroneous data:

```{r}
# temperature
dc2_subset %>% 
  ggplot(data = ., aes(x = date_time_PST, 
                       y = temp_C, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "red") + 
  geom_line(color = "red") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
# humdity
dc2_subset %>% 
  ggplot(data = ., aes(x = date_time_PST, 
                       y = relative_humidity, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "blue") + 
  geom_line(color = "blue") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-20")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-25")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-27")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-19")), 
             linetype = 4, color = "purple") +
  theme_classic()
```

That looks much better. However, HOBO "dry cool 2" is actually from the humid cool treatment. 


## HOBO "dry hot 1":

```{r}
# temperature
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Dry_Hot_1") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = temp_C, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "red") + 
  geom_line(color = "red") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-30")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
# humidity
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Dry_Hot_1") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = relative_humidity, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "blue") + 
  geom_line(color = "blue") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-30")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-11")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-12")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-19")), 
             linetype = 4, color = "purple") +
  theme_classic()
```


This logger was definitely in the (failed) hot treatments until April 30, so I don't want to use data from before then. Between April 30 and May 4, humidity is a little high. It also does not have the typical diel variation which is from the 12-12 light cycle, so the logger must have been out of the experiment. Likely, it was not placed in the cool treatment immediately after being removed from the hot, so I will exclude this period as well. I will also exclude the spikes at May 11, 12, and 19. 

Try excluding. For some reason the dates plotted versus filtered line up differently, which is why I filter out the date before what I plotted.

```{r}
unwanted_dates_dh1 <- as.Date(c(seq(as.Date("2021-04-19"), 
                                    as.Date("2021-05-03"), "days"),
                                "2021-05-10", "2021-05-11", "2021-05-18"),
                        format = "%Y-%m-%d")
dh1_subset <- format_HOBO_data %>%
  dplyr::filter(HOBO_ID == "Dry_Hot_1") %>%
  dplyr::filter(date_only %nin% unwanted_dates_dh1)
```

plot again to show that we removed erroneous data:

```{r}
# temperature
dh1_subset %>% 
  ggplot(data = ., aes(x = date_time_PST, 
                       y = temp_C, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "red") + 
  geom_line(color = "red") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
# humdity
dh1_subset %>% 
  ggplot(data = ., aes(x = date_time_PST, 
                       y = relative_humidity, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "blue") + 
  geom_line(color = "blue") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-11")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-12")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-19")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
```

Much more realistic. This is for the cool dry treatment. 


## HOBO "dry hot 2":

```{r}
# temperature
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Dry_Hot_2") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = temp_C, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "red") + 
  geom_line(color = "red") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-30")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-18")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-11")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-12")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
# humidity
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Dry_Hot_2") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = relative_humidity, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "blue") + 
  geom_line(color = "blue") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-18")), 
             linetype = 4, color = "purple") +
  theme_classic()
```

This logger was also in the hot treatment until April 30, and was also not put into a treatment until May 4. I'm not worried about May 11-12 so much for this one, it may have been out& back into a chamber within the logging interval. On May 18, however, the dip is a measurement from outside the chamber. 

Try excluding. For some reason the dates plotted versus filtered line up differently, which is why I filter out the date before what I plotted.

```{r}
unwanted_dates_dh2 <- as.Date(c(seq(as.Date("2021-04-19"), 
                                    as.Date("2021-05-03"), "days"),
                                "2021-05-17", "2021-05-18"),
                        format = "%Y-%m-%d")
dh2_subset <- format_HOBO_data %>%
  dplyr::filter(HOBO_ID == "Dry_Hot_2") %>%
  dplyr::filter(date_only %nin% unwanted_dates_dh2)
```

plot again to show that we removed erroneous data:

```{r}
# temperature
dh2_subset %>% 
  ggplot(data = ., aes(x = date_time_PST, 
                       y = temp_C, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "red") + 
  geom_line(color = "red") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-18")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
# humidity
dh2_subset %>% 
  ggplot(data = ., aes(x = date_time_PST, 
                       y = relative_humidity, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "blue") + 
  geom_line(color = "blue") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-18")), 
             linetype = 4, color = "purple") +
  theme_classic()
```


this logger was in cool & humid treatment

## HOBO "humid cool":

```{r}
# temperature
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Humid_Cool") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = temp_C, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "red") + 
  geom_line(color = "red") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
# humidity
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Humid_Cool") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = relative_humidity, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "blue") + 
  geom_line(color = "blue") +
  theme_classic()
```

This actually looks good as-is. We let the paper towel dry out around April 26, but that should be reflected. 

```{r}
hc_subset <- format_HOBO_data %>%
  dplyr::filter(HOBO_ID == "Humid_Cool")
```

And, the logger was named correctly. It was in humid and cool treatment. 


## HOBO "humid hot":

```{r}
# temperature
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Humid_Hot") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = temp_C, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "red") + 
  geom_line(color = "red") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-27")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-29")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-08")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-11")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-12")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
# humidity
format_HOBO_data %>% 
  dplyr::filter(HOBO_ID == "Humid_Hot") %>%
  ggplot(data = ., aes(x = date_time_PST, 
                       y = relative_humidity, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "blue") + 
  geom_line(color = "blue") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-27")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-04-29")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-04")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-08")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-11")), 
             linetype = 4, color = "purple") +
  geom_vline(xintercept = as.POSIXct(as.Date("2021-05-12")), 
             linetype = 4, color = "purple") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
```

This logger was in the hot treatment, then not in any treatment, same as the ones before, so I will exclude the data up to May 4. It looks fine to be used in the cool humid treatment average otherwise. 

```{r}
unwanted_dates_hh <- c(seq(as.Date("2021-04-19"), 
                                    as.Date("2021-05-03"), "days"))
hh_subset <- format_HOBO_data %>%
  dplyr::filter(HOBO_ID == "Humid_Hot") %>%
  dplyr::filter(date_only %nin% unwanted_dates_hh)
```

plot again to show that we removed erroneous data:

```{r}
# temperature
hh_subset %>% 
  ggplot(data = ., aes(x = date_time_PST, 
                       y = temp_C, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "red") + 
  geom_line(color = "red") +
  geom_hline(yintercept = 23) +
  geom_hline(yintercept = 29) +
  theme_classic()
# humidity
hh_subset %>% 
  ggplot(data = ., aes(x = date_time_PST, 
                       y = relative_humidity, 
                       color = HOBO_ID
                       )) +
  geom_point(color = "blue") + 
  geom_line(color = "blue") +
  theme_classic()
```




## Correct Tmt Groups & Trial Numbers

```{r}
tmts <- data.frame(HOBO_ID = unique(format_HOBO_data$HOBO_ID),
                   temp_tmt = as.factor(c("cool", "cool", "cool", 
                                "cool", "cool", "cool")),
                   humidity_tmt = as.factor(c("dry", "humid", "dry", 
                                    "humid", "humid", "humid")))
trials <- data.frame(trial_num = as.factor(c(rep("1", 8),
                                   rep("2", 7),
                                   rep("3", 7),
                                   rep("4", 8))),
                                    # trial 1
                     date_only = c(c(seq(as.Date("2021-04-19"), 
                                    as.Date("2021-04-26"), "days")),
                                   # trial 2
                                   c(seq(as.Date("2021-04-27"), 
                                    as.Date("2021-05-03"), "days")),
                                   # trial 3
                                   c(seq(as.Date("2021-05-04"), 
                                    as.Date("2021-05-10"), "days")),
                                   # trial 4
                                   c(seq(as.Date("2021-05-11"), 
                                    as.Date("2021-05-18"), "days"))
                                  )
                     )
```


## Rejoin All

```{r}
clean_HOBO_data <- dc1_subset %>%
  rbind(dc2_subset) %>%
  rbind(dh1_subset) %>%
  rbind(dh2_subset) %>%
  rbind(hc_subset) %>%
  rbind(hh_subset) %>%
  left_join(tmts, by = 'HOBO_ID') %>%
  left_join(trials, by = 'date_only')
summary(clean_HOBO_data)
```



# Get Means

## Grand Mean

```{r}
grand_mean <- clean_HOBO_data %>%
  group_by(temp_tmt) %>%
  summarise(temp_mean = mean(temp_C),
            temp_SD = sd(temp_C))
grand_mean
```

## Treatment Overall

```{r}
tmt_only_means <- clean_HOBO_data %>%
  group_by(humidity_tmt) %>%
  summarise(temp_mean = mean(temp_C),
            temp_SD = sd(temp_C),
            humidity_mean = mean(relative_humidity),
            humidity_SD = sd(relative_humidity))
tmt_only_means
```



## Treatment by Trial

```{r}
tmt_trial_means <- clean_HOBO_data %>%
  group_by(humidity_tmt, trial_num) %>%
  summarise(temp_mean = mean(temp_C),
            temp_SD = sd(temp_C),
            humidity_mean = mean(relative_humidity),
            humidity_SD = sd(relative_humidity)) %>%
  arrange(trial_num)
tmt_trial_means
write.csv(tmt_trial_means, "./HOBO_tmt_trial_diffs.csv")
```



## Check Statistical Differences

```{r}
# based on humidity treatment
temp_tmt_mod <- lm(data = clean_HOBO_data, 
               temp_C ~ humidity_tmt)
summary(temp_tmt_mod)
humidity_tmt_mod <- lm(data = clean_HOBO_data, 
                   relative_humidity ~ humidity_tmt)
summary(humidity_tmt_mod)
# based on trial
temp_trial_mod <- lm(data = clean_HOBO_data, 
               temp_C ~ trial_num)
summary(temp_trial_mod)
humidity_trial_mod <- lm(data = clean_HOBO_data, 
                   relative_humidity ~ trial_num)
summary(humidity_trial_mod)
```

temp summary: dry mean = 25.1, humid mean = 24.9, estimate = -0.225, t = -5.069, p < 0.0001

humidity summary: estimate = 69.232, t = 264.61, p < 0.0001


Both humidity treatment and trial number, AND their interaction is significant for predicting the temperature and humidity the data loggers logged. 

Need pairwise to know which ones are actually different:

```{r}
# humidity 
anova_humidity <- aov(humidity_mod)
posthoc_humidity <- TukeyHSD(anova_humidity)
posthoc_humidity
# temp
anova_temp <- aov(temp_mod)
posthoc_temp <- TukeyHSD(anova_temp)
posthoc_temp
```
