---
title: "Climate Water Loss Experiment - Plasma Osmolality Data Wrangling"
author: "Savannah Weaver"
date: "2021"
output: pdf_document
toc: TRUE
---


# Packages

```{r setup, include = TRUE}
`%nin%` = Negate(`%in%`)
if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse") # workflow and plots
```


# Background and Goals

Blood was drawn from the postorbital sinus of adult male *Sceloporus occidentalis* between June - August 2021. After centrifuging and separating, plasma was run on a VAPRO vapor pressure osmometer in 1-4 replicates, when plasma volume allowed. In this R script, I check the distribution of replicates, omit outliers, and average remaining replicates. The final values will be more precise and accurate estimates of the true plasma osmolality for each lizard, and those values will be used in the capture_analysis and experiment_analysis R script files. Please refer to **doi:** for the published scientific journal article and full details.


# Load Data

```{r load data}
osml_reps <- read.csv("./data/osmolality.csv",
                na.strings = c("","NA"),
                header = TRUE
                ) %>%
    dplyr::mutate(date_blood_drawn = as.Date(date_blood_drawn,
                                             format = "%m/%d/%y"),
                  date_osmom_run = as.Date(date_osmom_run,
                                           format = "%m/%d/%y"),
                  time_osmom_run = as.POSIXct(time_osmom_run, 
                                               format = "%H:%M"),
                  individual_ID = as.factor(individual_ID),
                  replicate_no = as.factor(replicate_no),
                  osmolality_mmol_kg = as.numeric(osmolality_mmol_kg)
                  )
summary(osml_reps)
```


# Check Data


## Dates

Blood was drawn on day 0, 4, 8, and 10 of the experiment. Create a list of the dates expected to have blood draw data, then determine whether I have dates outside those.

Trail 1: June 16-24
Trail 2: June 26 - July 4
Trial 3: July 20-28
Trial 4: August 8-16
Trial 5: August 22-30

```{r check dates}
                            # trial 1
expected_dates <- as.Date(c("2021-06-16", "2021-06-20", 
                            "2021-06-24", "2021-06-26",
                            # trial 2
                            "2021-06-26", "2021-06-30", 
                            "2021-07-04", "2021-07-06",
                            # trial 3
                            "2021-07-20", "2021-07-24", 
                            "2021-07-28", "2021-07-30",
                            # trial 4
                            "2021-08-08", "2021-08-12", 
                            "2021-08-16", "2021-08-18",
                            # trial 5
                            "2021-08-22", "2021-08-26", 
                            "2021-08-30", "2021-09-01"))

length(osml_reps$date_blood_drawn[osml_reps$date_blood_drawn %nin% expected_dates]
)
```

There are zero blood draw dates that are not in our expected list.


## Number of Blood Draws

Each lizard should have had their blood drawn on 4 different dates, unless they were taken out of the experiment early.

```{r check n obs}
# get ID's of the individuals that completed treatment
individuals <- read.csv("./data/tmt_assignments.csv") %>%
  dplyr::select(conclusion, individual_ID) %>%
  dplyr::filter(conclusion == "complete") %>%
  mutate(individual_ID = as.factor(individual_ID),
         conclusion = as.factor(conclusion))
summary(individuals)

# calculate the number of dates for each individual
osml_reps %>%
  dplyr::filter(individual_ID %in% individuals$individual_ID) %>%
  group_by(individual_ID, date_blood_drawn) %>%
  summarise(n()) %>%
  group_by(individual_ID) %>%
  summarise(n()) %>%
  arrange(n())
```


Wahoo, every lizard has blood draws from 4 different dates.






# Replicates

Now, I will try to identify outliers within the replicates for a given individual on a given date. There must be at least 3 replicates to do this, so the first thing I need to do is figure out which individuals/dates have enough replicates, then subset my data to be only those individuals.


## Individuals w 3+ Replicates

```{r separate data w enough reps to assess}
# identify individuals with 3-4 reps
enuf_reps <- osml_reps %>%
  group_by(individual_ID, date_blood_drawn) %>%
  mutate(count = n()) %>%
  dplyr::filter(count > 2) %>%
  arrange(count)
enuf_reps

# identify individuals with 1-2 reps
not_reps <- osml_reps %>%
  group_by(individual_ID, date_blood_drawn) %>%
  mutate(count = n()) %>%
  dplyr::filter(count < 3) %>%
  arrange(count)
not_reps

# check total obs still add to original 1484
nrow(enuf_reps) + nrow(not_reps)
nrow(enuf_reps) + nrow(not_reps) == nrow(osml_reps)
```


## Assess Variation

We want the Coefficient of Variation (CV) among our technical replicates to be small. We need to calculate it to identify whether there may be outliers.

```{r calculate CVs}
CVs <- enuf_reps %>%
  group_by(individual_ID, date_blood_drawn) %>%
  summarise(mean = mean(osmolality_mmol_kg),
            SD = sd(osmolality_mmol_kg),
            CV = (SD/mean) *100,
            min = min(osmolality_mmol_kg),
            max = max(osmolality_mmol_kg),
            range = max - min
            )
summary(CVs)
hist(CVs$CV)
hist(CVs$range) 
```

Ideally, CV would be 10-15%. If it's larger, and one of the replicates is very different than the others, we can assume that the replicates that are closer together are more reliable. 

The CV is >10 for only one lizard on one date, so our replicates are already likely to accurately represent the true value. We don't want to lose accuracy by searching for precision, so we will only remove the one point driving the enormous CV value.


## Find Outlier

```{r create find outlier function}
# write function to find outliers for each individual on each date
find_outliers <- function(df) {
  
  # initiate dataframe to compile info and list to compile plots
  outliers <- data.frame()
  #boxplots <- list()

  # initiate a for loop to go through every who in df
  for(indiv_ch in unique(df$individual_ID)) {
    
    # select data for only the individual of interest
    df_sub <- df %>%
      dplyr::filter(individual_ID == as.numeric(indiv_ch))
    
    # make a boxplot
    df_sub %>%
      ggplot(.) +
      geom_boxplot(aes(x = as.factor(date_blood_drawn),
                       y = osmolality_mmol_kg,
                       fill = as.factor(date_blood_drawn))) +
      ggtitle(paste("Individual", indiv_ch)) +
      theme_classic() -> plot
    
    # print/save
    print(plot)
    #boxplots[[indiv_ch]] <- plot
    
    # extract outliers
    outs <- df_sub %>%
      group_by(individual_ID, date_blood_drawn) %>%
      summarise(outs = boxplot.stats(osmolality_mmol_kg)$out)
    
    # add to running dataframe of outliers
    outliers <- outliers %>%
      rbind(outs)
  }
  #return(boxplots)
  return(outliers)
}
```


Now apply the function to the data:

```{r find outlier using boxplots, fig.show="hold", out.width="50%"}
par(mfrow = c(71, 2))
outliers_found <- find_outliers(enuf_reps)
outliers_found
par(mfrow = c(1, 1))
```

Based on boxplots, there are no outliers, but we still want to find and omit the point severely increasing CV for one individual on one date.



Determine which replicate is an outlier and pulls the CV above our acceptable range (10-15%).

```{r find outlier using CV}
CV_all <- enuf_reps %>%
    group_by(individual_ID, date_blood_drawn) %>%
    summarise(mean = mean(osmolality_mmol_kg),
              SD = sd(osmolality_mmol_kg),
              CV_all = (SD/mean) *100) %>%
    dplyr::select(-mean, -SD)
  
CV_12 <- enuf_reps %>%
    dplyr::filter(replicate_no != 3) %>%
    group_by(individual_ID, date_blood_drawn) %>%
    summarise(mean = mean(osmolality_mmol_kg),
              SD = sd(osmolality_mmol_kg),
              CV_12 = (SD/mean) *100) %>%
    dplyr::select(-mean, -SD)
  
CV_23 <- enuf_reps %>%
    dplyr::filter(replicate_no != 1) %>%
    group_by(individual_ID, date_blood_drawn) %>%
    summarise(mean = mean(osmolality_mmol_kg),
              SD = sd(osmolality_mmol_kg),
              CV_23 = (SD/mean) *100) %>%
    dplyr::select(-mean, -SD)
  
CV_31 <- enuf_reps %>%
    dplyr::filter(replicate_no != 2) %>%
    group_by(individual_ID, date_blood_drawn) %>%
    summarise(mean = mean(osmolality_mmol_kg),
              SD = sd(osmolality_mmol_kg),
              CV_31 = (SD/mean) *100) %>%
    dplyr::select(-mean, -SD)
  
compare <- CV_all %>%
    left_join(CV_12, by = c('individual_ID', 'date_blood_drawn')) %>%
    left_join(CV_23, by = c('individual_ID', 'date_blood_drawn')) %>%
    left_join(CV_31, by = c('individual_ID', 'date_blood_drawn')) %>%
  arrange(desc(CV_all)) %>%
  dplyr::filter(CV_all>10)
compare
```

We should remove replicate 2 for individual 225 on June 26.


## Remove Outlier

```{r remove outlier}
enuf_reps[199:201, 5:6]
enuf_reps[200, "osmolality_mmol_kg"] = NA 
enuf_reps[199:201, 5:6]

# check
enuf_reps %>%
  dplyr::filter(individual_ID == 225,
                date_blood_drawn == as.Date("2021-06-26", 
                                              format= "%Y-%m-%d")
                #replicate_no == 2
                )

# filter out the one NA
enuf_reps_trimmed <- enuf_reps %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg))
```



## Average Remaining Replicates

Now that the outliers are removed from the technical replicates when there were enough replicates to identify them, I will average the remaining replicates of the rejoined data for lizards with 1-2 and 3-4 replicates.

```{r calculate means}
osml_means <- not_reps %>%
  rbind(enuf_reps_trimmed) %>%
  group_by(date_blood_drawn, individual_ID) %>%
  summarise(osmolality_mmol_kg_mean = mean(osmolality_mmol_kg))
```


# Export

```{r export cleaned}
write.csv(osml_means, "./data/osml_means_clean.csv")
```









